---
title: "Abundance Trends"
output: html_document
date: "2023-07-03"
---

##Libraries

```{r setup, include=FALSE}
library(readxl)
library(here)
library(dplyr)
library(tidyverse)
library(ggplot2)
library(lubridate)
library(data.table)
library(chron)
```

##Data and Removing Duplicates

```{r}
dat <- read_xlsx(here::here("Data", "AbundanceData.xlsx"))
removed_dat <-  read_xlsx(here::here("Data", "AbundanceData.xlsx"))

#Removing duplicates of events
removed_dat$Date <- as.Date(removed_dat$Date)
duplicates <- duplicated(removed_dat$TimeBegin) & duplicated(removed_dat$Date)
unique_dat <- removed_dat[!duplicates, ]

setDT(unique_dat)
unique_dat[, TimeBegin := as.ITime(sprintf("%s", TimeBegin), format = "%H%M")]
unique_dat[, TimeEnd := as.ITime(sprintf("%s", TimeEnd), format = "%H%M")]
```


##Time dataframe 

```{r}
time_dat <- dat %>%
  filter(!is.na(Time)) %>%
  group_by(Time, Year) %>%
  summarize(Sum_Turtles = sum(Num_Turtles)) %>%
  summarize(Mean_Turtles = mean(Sum_Turtles))
  

time_dat$Time <- as.numeric(time_dat$Time)
time_dat$Time <- sprintf("%02d:%02d", time_dat$Time %/% 100, time_dat$Time %% 100)
time_dat$Time <- as.POSIXct(time_dat$Time, format = "%H:%M")

```

##Avg start and end times for each island

Finding the average start and end time for surveys at each island

```{r}
#Finding the average survey times for Laysan
Laysan_dat <- unique_dat %>%
  filter(Site == "LAY")

Laysan_avg_begin_time_numeric <- mean(as.numeric(Laysan_dat$TimeBegin), na.rm = TRUE)
Laysan_avg_begin_time <- format(as.ITime(Laysan_avg_begin_time_numeric, format = "h:m:s"), format = "%H:%M")
Laysan_avg_begin_time

Laysan_avg_end_time_numeric <- mean(as.numeric(Laysan_dat$TimeEnd), na.rm = TRUE)
Laysan_avg_end_time <- format(as.ITime(Laysan_avg_end_time_numeric, format = "h:m:s"), format = "%H:%M")
Laysan_avg_end_time

#Finding the average survey times for Lisianski
Lisi_dat <- unique_dat %>%
  filter(Site == "LIS")

Lisi_avg_begin_time_numeric <- mean(as.numeric(Lisi_dat$TimeBegin), na.rm = TRUE)
Lisi_avg_begin_time <- format(as.ITime(Lisi_avg_begin_time_numeric, format = "h:m:s"), format = "%H:%M")
Lisi_avg_begin_time

Lisi_avg_end_time_numeric <- mean(as.numeric(Lisi_dat$TimeEnd), na.rm = TRUE)
Lisi_avg_end_time <- format(as.ITime(Lisi_avg_end_time_numeric, format = "h:m:s"), format = "%H:%M")
Lisi_avg_end_time

#Finding the average survey times for Midway
Midway_dat <- unique_dat %>%
  filter(Site == "MDY")

Midway_avg_begin_time_numeric <- mean(as.numeric(Midway_dat$TimeBegin), na.rm = TRUE)
Midway_avg_begin_time <- format(as.ITime(Midway_avg_begin_time_numeric, format = "h:m:s"), format = "%H:%M")
Midway_avg_begin_time

Midway_avg_end_time_numeric <- mean(as.numeric(Midway_dat$TimeEnd), na.rm = TRUE)
Midway_avg_end_time <- format(as.ITime(Midway_avg_end_time_numeric, format = "h:m:s"), format = "%H:%M")
Midway_avg_end_time

#Finding the average survey times for Pearl & Hermes
PH_dat <- unique_dat %>%
  filter(Site == "PHR")

PH_avg_begin_time_numeric <- mean(as.numeric(PH_dat$TimeBegin), na.rm = TRUE)
PH_avg_begin_time <- format(as.ITime(PH_avg_begin_time_numeric, format = "h:m:s"), format = "%H:%M")
PH_avg_begin_time

PH_avg_end_time_numeric <- mean(as.numeric(PH_dat$TimeEnd), na.rm = TRUE)
PH_avg_end_time <- format(as.ITime(PH_avg_end_time_numeric, format = "h:m:s"), format = "%H:%M")
PH_avg_end_time

#Finding the average survey times for Kure
Kure_dat <- unique_dat %>%
  filter(Site == "KUR")

Kure_avg_begin_time_numeric <- mean(as.numeric(Kure_dat$TimeBegin), na.rm = TRUE)
Kure_avg_begin_time <- format(as.ITime(Kure_avg_begin_time_numeric, format = "h:m:s"), format = "%H:%M")
Kure_avg_begin_time

Kure_avg_end_time_numeric <- mean(as.numeric(Kure_dat$TimeEnd), na.rm = TRUE)
Kure_avg_end_time <- format(as.ITime(Kure_avg_end_time_numeric, format = "h:m:s"), format = "%H:%M")
Kure_avg_end_time

```
##Graph for Time trends

This graph is a plot of the average of turtles seen at each time of day. It is combining all Northwest Islands and all years. The colored bars represent the average start and end time for surveys done at each island across all years. Blue is Laysan, red is Laysan, green is Midway, orange is Pearl and Hermes and yellow is Kure. 

```{r}

time_trend <- ggplot() +
  geom_line(data = time_dat, aes(x = Time, y = Mean_Turtles)) +
  labs(x = "Time", y = "Number of Honu Seen", title = "Time of Honu Seen During Basking Surveys (1983-2022)", color = "Legend") +
  scale_x_datetime(name = "Time", 
                   breaks = seq(as.POSIXct("07:00", format = "%H:%M"), 
                                as.POSIXct("20:00", format = "%H:%M"), 
                                by = "1 hour"), 
                   date_labels = "%H:%M", expand = c(0,0)) +
  theme_minimal() +
  annotate("rect", xmin = as.POSIXct("2023-07-17 12:59:44"), xmax = as.POSIXct("2023-07-17 15:38:08"), ymin = 0, ymax = 2,
           alpha = .3, fill = "blue") +
  annotate("rect", xmin = as.POSIXct("2023-07-17 13:07:00"), xmax = as.POSIXct("2023-07-17 14:59:51"), ymin = 2, ymax = 4,
           alpha = .3, fill = "red") +
  annotate("rect", xmin = as.POSIXct("2023-07-17 12:37:46"), xmax = as.POSIXct("2023-07-17 15:10:25"), ymin = 4, ymax = 6,
           alpha = .3, fill = "green") +
  annotate("rect", xmin = as.POSIXct("2023-07-17 13:16:45"), xmax = as.POSIXct("2023-07-17 14:08:51"), ymin = 6, ymax = 8,
           alpha = .3, fill = "orange") +
  annotate("rect", xmin = as.POSIXct("2023-07-17 13:47:45"), xmax = as.POSIXct("2023-07-17 15:09:01"), ymin = 8, ymax = 10,
           alpha = .3, fill = "yellow") +
  scale_fill_manual(name = "Island", values = c("Laysan" = "blue", "Lisianski" = "red", "Midway" = "green", "Pearl and Hermes" = "orange", "Kure" = "yellow"), labels = c("Laysan", "Lisianski", "Midway", "Pearl and Hermes", "Kure")) +
  guides(fill = guide_legend(title = "Island")) +
  theme(panel.border = element_rect(color = "black", fill = NA)) +
  scale_y_continuous(limits = c(0, 10), expand = c(0,0), breaks = seq(0, 10, by = 2))

time_trend 
```

##Time Dataframes for each island

```{r}
time_dat_site <- dat %>%
  filter(!is.na(Time)) %>%
  group_by(Time, Year, Site) %>%
  summarize(Sum_Turtles = sum(Num_Turtles))

time_dat_site <- time_dat_site %>%
  group_by(Time, Site) %>%
  summarize(Mean_Turtles = mean(Sum_Turtles))

time_dat_site$Time <- sprintf("%04d", as.integer(time_dat_site$Time))
time_dat_site$Time <- strptime(time_dat_site$Time, format = "%H%M")
time_dat_site$Time <- as.POSIXct(time_dat_site$Time)

Laysan_time_dat <- time_dat_site %>%
  filter(Site == "LAY")

Lisi_time_dat <- time_dat_site %>%
  filter(Site == "LIS")

Midway_time_dat <- time_dat_site %>%
  filter(Site == "MDY")

PH_time_dat <- time_dat_site %>%
  filter(Site == "PHR")

Kure_time_dat <- time_dat_site %>%
  filter(Site == "KUR")
```

##Laysan time graph

```{r}
Laysan_time <- ggplot() +
  geom_line(data = Laysan_time_dat, aes(x = Time, y = Mean_Turtles)) +
  labs(x = "Time", y = "Number of Honu Seen", title = "Time of Honu Seen During Basking Surveys On Kamole (1983-2022, n = 34)", color = "Legend") +
  scale_x_datetime(name = "Time", 
                   breaks = seq(as.POSIXct("07:00", format = "%H:%M"), 
                                as.POSIXct("21:00", format = "%H:%M"), 
                                by = "1 hour"), limits = c(as.POSIXct("07:00", format = "%H:%M"), 
                                as.POSIXct("21:00", format = "%H:%M")), 
                   date_labels = "%H:%M") +
  theme_minimal() +
  annotate("rect", xmin = as.POSIXct("2023-07-18 12:59:44"), xmax = as.POSIXct("2023-07-18 15:38:08"), ymin = 1, ymax = 4,
           alpha = .3, fill = "blue") +
  theme(panel.border = element_rect(color = "black", fill = NA), plot.margin = margin(5.5, 10, 5.5, 5.5)) 

Laysan_time
```
##Lisianski Time Graph

```{r}
Lisi_time <- ggplot() +
  geom_line(data = Lisi_time_dat, aes(x = Time, y = Mean_Turtles)) +
  labs(x = "Time", y = "Number of Honu Seen", title = "Time of Honu Seen During Basking Surveys On Kapou (1983-2022, n = 32)", color = "Legend") +
  scale_x_datetime(name = "Time", 
                   breaks = seq(as.POSIXct("07:00", format = "%H:%M"), 
                                as.POSIXct("21:00", format = "%H:%M"), 
                                by = "1 hour"), limits = c(as.POSIXct("07:00", format = "%H:%M"), 
                                as.POSIXct("21:00", format = "%H:%M")), 
                   date_labels = "%H:%M") +
  theme_minimal()+ 
  annotate("rect", xmin = as.POSIXct("2023-07-18 13:07:00"), xmax = as.POSIXct("2023-07-18 14:59:51"), ymin = 1, ymax = 6,
           alpha = .3, fill = "red") +
  theme(panel.border = element_rect(color = "black", fill = NA), plot.margin = margin(5.5, 10, 5.5, 5.5))

Lisi_time
```

##Midway Time Graph

```{r}
Midway_time <- ggplot() +
  geom_line(data = Midway_time_dat, aes(x = Time, y = Mean_Turtles)) +
  labs(x = "Time", y = "Number of Honu Seen", title = "Time of Honu Seen During Basking Surveys On Midway (1992-2022, n = 23)", color = "Legend") +
  scale_x_datetime(name = "Time", 
                   breaks = seq(as.POSIXct("07:00", format = "%H:%M"), 
                                as.POSIXct("21:00", format = "%H:%M"), 
                                by = "1 hour"), limits = c(as.POSIXct("07:00", format = "%H:%M"), 
                                as.POSIXct("21:00", format = "%H:%M")), 
                   date_labels = "%H:%M") +
  theme_minimal()+
  annotate("rect", xmin = as.POSIXct("2023-07-18 12:37:46"), xmax = as.POSIXct("2023-07-18 15:10:25"), ymin = 1, ymax = 7,
           alpha = .3, fill = "green") +
  theme(panel.border = element_rect(color = "black", fill = NA), plot.margin = margin(5.5, 10, 5.5, 5.5))

Midway_time
```

##Pearl and Hermes Time Graph

```{r}
PH_time <- ggplot() +
  geom_line(data = PH_time_dat, aes(x = Time, y = Mean_Turtles)) +
  labs(x = "Time", y = "Number of Honu Seen", title = "Time of Honu Seen During Basking Surveys On Manawai (1983-2022, n = 34)", color = "Legend") +
  scale_x_datetime(name = "Time", 
                   breaks = seq(as.POSIXct("07:00", format = "%H:%M"), 
                                as.POSIXct("21:00", format = "%H:%M"), 
                                by = "1 hour"), limits = c(as.POSIXct("07:00", format = "%H:%M"), 
                                as.POSIXct("21:00", format = "%H:%M")), 
                   date_labels = "%H:%M") +
  theme_minimal()+
  annotate("rect", xmin = as.POSIXct("2023-07-18 13:16:45"), xmax = as.POSIXct("2023-07-18 14:08:51"), ymin = 1, ymax = 16,
           alpha = .3, fill = "orange") +
  theme(panel.border = element_rect(color = "black", fill = NA), plot.margin = margin(5.5, 10, 5.5, 5.5))

PH_time
```

##Kure Time graphs

```{r}
Kure_time <- ggplot() +
  geom_line(data = Kure_time_dat, aes(x = Time, y = Mean_Turtles)) +
  labs(x = "Time", y = "Number of Honu Seen", title = "Time of Honu Seen During Basking Surveys On Holaniku (1983-2022, n = 32)", color = "Legend") +
  scale_x_datetime(name = "Time", 
                   breaks = seq(as.POSIXct("07:00", format = "%H:%M"), 
                                as.POSIXct("21:00", format = "%H:%M"), 
                                by = "1 hour"), limits = c(as.POSIXct("07:00", format = "%H:%M"), 
                                as.POSIXct("21:00", format = "%H:%M")), 
                   date_labels = "%H:%M") +
  theme_minimal()+
  annotate("rect", xmin = as.POSIXct("2023-07-18 13:47:45"), xmax = as.POSIXct("2023-07-18 15:09:01"), ymin = 1, ymax = 3,
           alpha = .3, fill = "yellow") +
  theme(panel.border = element_rect(color = "black", fill = NA), plot.margin = margin(5.5, 10, 5.5, 5.5))

Kure_time
```

##Month Data Means


```{r}
month_dat <- dat %>%
  group_by(Date, Year) %>%
  summarize(Sum_Turtles = sum(Num_Turtles))
month_dat$MonthDay <- format(month_dat$Date, "%m-%d") %>%
  na.omit()

month_dat <- month_dat %>%
  group_by(MonthDay) %>%
  summarize(Mean_Turtles = mean(Sum_Turtles))
month_dat$MonthDay <- as.Date(month_dat$MonthDay, format = "%m-%d")


 first_days <- (format(month_dat$MonthDay, "%m-01"))
```


##Finding average start and end survey dates at each island

```{r}
avg_month <- dat %>%
  group_by(Date, Year, Site) %>%
  summarize(Sum_Turtles = sum(Num_Turtles))
avg_month$MonthDay <- format(avg_month$Date, "%m-%d")

#Laysan avg date
Laysan_starts <- avg_month %>%
  filter(Site == "LAY") %>%
  group_by(Year) %>%
  filter(Date == min(Date)) 
Laysan_starts$MonthDay <- as.Date(Laysan_starts$MonthDay, format = "%m-%d")

Laysan_ends <- avg_month %>%
  filter(Site == "LAY") %>%
  group_by(Year) %>%
  filter(Date == max(Date))
Laysan_ends$MonthDay <- as.Date(Laysan_ends$MonthDay, format = "%m-%d")
  
avg_start_L <- mean(Laysan_starts$MonthDay)
avg_ends_L <- mean(Laysan_ends$MonthDay)

avg_start_L
avg_ends_L

#Lisianski avg values
Lisi_starts <- avg_month %>%
  filter(Site == "LIS") %>%
  group_by(Year) %>%
  filter(Date == min(Date)) 
Lisi_starts$MonthDay <- as.Date(Lisi_starts$MonthDay, format = "%m-%d")

Lisi_ends <- avg_month %>%
  filter(Site == "LIS") %>%
  group_by(Year) %>%
  filter(Date == max(Date))
Lisi_ends$MonthDay <- as.Date(Lisi_ends$MonthDay, format = "%m-%d")
  
avg_start_Li <- mean(Lisi_starts$MonthDay)
avg_ends_Li <- mean(Lisi_ends$MonthDay)

avg_start_Li
avg_ends_Li

#Midway avg dates
Midway_starts <- avg_month %>%
  filter(Site == "MDY") %>%
  group_by(Year) %>%
  filter(Date == min(Date)) 
Midway_starts$MonthDay <- as.Date(Midway_starts$MonthDay, format = "%m-%d")

Midway_ends <- avg_month %>%
  filter(Site == "MDY") %>%
  group_by(Year) %>%
  filter(Date == max(Date))
Midway_ends$MonthDay <- as.Date(Midway_ends$MonthDay, format = "%m-%d")
  
avg_start_M <- mean(Midway_starts$MonthDay)
avg_ends_M <- mean(Midway_ends$MonthDay)

avg_start_M
avg_ends_M

#PH avg dates
PH_starts <- avg_month %>%
  filter(Site == "PHR") %>%
  group_by(Year) %>%
  filter(Date == min(Date)) 
PH_starts$MonthDay <- as.Date(PH_starts$MonthDay, format = "%m-%d")

PH_ends <- avg_month %>%
  filter(Site == "PHR") %>%
  group_by(Year) %>%
  filter(Date == max(Date))
PH_ends$MonthDay <- as.Date(PH_ends$MonthDay, format = "%m-%d")
  
avg_start_P <- mean(PH_starts$MonthDay)
avg_ends_P <- mean(PH_ends$MonthDay)

avg_start_P
avg_ends_P

#Kure avg dates
Kure_starts <- avg_month %>%
  filter(Site == "KUR") %>%
  group_by(Year) %>%
  filter(Date == min(Date)) 
Kure_starts$MonthDay <- as.Date(Kure_starts$MonthDay, format = "%m-%d")

Kure_ends <- avg_month %>%
  filter(Site == "KUR") %>%
  group_by(Year) %>%
  filter(Date == max(Date))
Kure_ends$MonthDay <- as.Date(Kure_ends$MonthDay, format = "%m-%d")
  
avg_start_K <- mean(Kure_starts$MonthDay)
avg_ends_K <- mean(Kure_ends$MonthDay)

avg_start_K
avg_ends_K
```

##Month Trend totals graph

This graph is showing the average of turtles seen in one day on every island and across every year. This is coming from all the complete censuses that we have from the Monk Seal Program and the MTBAP program from 1983 to 2022. The colored bars represent the average range of surveys done at each island. I found the average earliest and latest survey done each year at each island. Blue is Laysan, red is Laysan, green is Midway, orange is Pearl and Hermes and yellow is Kure. 

```{r}
month_trend <- ggplot(month_dat, aes(x = MonthDay, y = Mean_Turtles)) +
  geom_line() +
  scale_x_date(breaks = seq(as.Date("2023-01-01"), as.Date("2023-12-31"), by = "month"), date_labels = "%b", limits = c(as.Date("2023-01-01"), as.Date("2023-12-31")), expand = c(0.01,0))+
  labs(x = "Month", y = "Number of Honu Seen", title = "Average Number of Honu Seen Each Day (1983-2022)") +
  theme_minimal() +
  annotate("rect", xmin = as.Date("2023-04-24"), xmax = as.Date("2023-07-29"), ymin = 0, ymax = 4,
           alpha = .3, fill = "blue") +
  annotate("rect", xmin = as.Date("2023-05-21"), xmax = as.Date("2023-07-28"), ymin = 4, ymax = 8,
           alpha = .3, fill = "red") +
  annotate("rect", xmin = as.Date("2023-04-09"), xmax = as.Date("2023-08-26"), ymin = 8, ymax = 12,
           alpha = .3, fill = "green") +
  annotate("rect", xmin = as.Date("2023-06-18"), xmax = as.Date("2023-07-29"), ymin = 12, ymax = 16,
           alpha = .3, fill = "orange") +
  annotate("rect", xmin = as.Date("2023-05-25"), xmax = as.Date("2023-07-23"), ymin = 16, ymax = 20,
           alpha = .3, fill = "yellow") +
  theme(panel.border = element_rect(color = "black", fill = NA)) +
  scale_y_continuous(limits = c(0, 20), expand = c(0,0), breaks = seq(0, 20, by = 4))

month_trend
```
##Dataframe for individual islands with averages

```{r}
month_dat_site <- dat %>%
  group_by(Date, Year, Site) %>%
  summarize(Sum_Turtles = sum(Num_Turtles))
month_dat_site$MonthDay <- format(month_dat_site$Date, "%m-%d") %>%
  na.omit()

month_dat_site <- month_dat_site %>%
  group_by(MonthDay, Site) %>%
  summarize(Mean_Turtles = mean(Sum_Turtles))
month_dat_site$MonthDay <- as.Date(month_dat_site$MonthDay, format = "%m-%d")
```



##Laysan month graph

The trend line graph is the same thing as the above graph showing turtles seen by month, but just for Laysan. The colored block is to represent the average earliest and latest survey date across all years. The graph is repeated with data from each of the other islands.

```{r}
Laysan_dat2 <- month_dat_site %>%
  filter(Site == "LAY")
Laysan_dat2$MonthDay <- as.Date(Laysan_dat2$MonthDay, format = "%m-%d")


Laysan_month <- ggplot(data = Laysan_dat2, aes(x = MonthDay, y = Mean_Turtles)) +
  geom_line() +
  scale_x_date(breaks = seq(as.Date("2023-01-01"), as.Date("2023-12-31"), by = "month"), date_labels = "%b", limits = c(as.Date("2023-01-01"), as.Date("2023-12-31")), expand = c(0.01,0))+
  labs(x = "Month", y = "Number of Honu Seen", title = "Average Number of Honu Seen by Month at Kamole (1983-2022, n = 34)") +
  theme_minimal() +
  theme(plot.margin = margin(5.5, -50, 5.5, 5.5)) +
  annotate("rect", xmin = as.Date("2023-04-24"), xmax = as.Date("2023-07-29"), ymin = 0, ymax = 15,
           alpha = .3, fill = "blue") +
  geom_smooth(data = Laysan_dat2, linetype = "dashed", color = "black", se = FALSE, method = 'loess') +
  theme(panel.border = element_rect(color = "black", fill = NA), plot.margin = margin(5.5, 10, 5.5, 5.5)) +
  scale_y_continuous(limits = c(0,15), expand = c(0,0))

Laysan_month
```

##Lisi month graph

```{r}
Lisi_dat2 <- month_dat_site %>%
  filter(Site == "LIS")
Lisi_dat2$MonthDay <- as.Date(Lisi_dat2$MonthDay, format = "%m-%d")

Lisi_month <- ggplot(data = Lisi_dat2, aes(x = MonthDay, y = Mean_Turtles)) +
  geom_line() +
 scale_x_date(breaks = seq(as.Date("2023-01-01"), as.Date("2023-12-31"), by = "month"), date_labels = "%b", limits = c(as.Date("2023-01-01"), as.Date("2023-12-31")), expand = c(0.01,0)) +
  labs(x = "Month", y = "Number of Honu Seen", title = "Average Number of Honu Seen by Month at Kapou (1983-2022, n = 32)") +
  theme_minimal() +
  theme(plot.margin = margin(5.5, 10, 5.5, 5.5)) +
  annotate("rect", xmin = as.Date("2023-05-21"), xmax = as.Date("2023-07-28"), ymin = 0, ymax = 40,
           alpha = .3, fill = "red") +
  geom_smooth(data = Lisi_dat2, linetype = "dashed", color = "black", se = FALSE, method = 'loess') +
  theme(panel.border = element_rect(color = "black", fill = NA)) +
  scale_y_continuous(limits = c(0, 40), expand = c(0,0))

Lisi_month
```

##Midway month graph

```{r}
Midway_dat2 <- month_dat_site %>%
  filter(Site == "MDY")
Midway_dat2$MonthDay <- as.Date(Midway_dat2$MonthDay, format = "%m-%d")

Midway_month <- ggplot(data =Midway_dat2, aes(x = MonthDay, y = Mean_Turtles)) +
  geom_line() +
  scale_x_date(breaks = seq(as.Date("2023-01-01"), as.Date("2023-12-31"), by = "month"), date_labels = "%b", limits = c(as.Date("2023-01-01"), as.Date("2023-12-31")), expand = c(0.01,0))+
  labs(x = "Month", y = "Number of Honu Seen", title = "Average Number of Honu Seen by Month at Midway (1992-2022, n = 23)") +
  theme_minimal() +
  theme(plot.margin = margin(5.5, 10, 5.5, 5.5)) +
  annotate("rect", xmin = as.Date("2023-04-09"), xmax = as.Date("2023-08-26"), ymin = 0, ymax = 30,
           alpha = .3, fill = "green") +
  geom_smooth(data = Midway_dat2, linetype = "dashed", color = "black", se = FALSE, method = 'loess') +
  theme(panel.border = element_rect(color = "black", fill = NA), plot.margin = margin(5.5, 10, 5.5, 5.5)) +
  scale_y_continuous(limits = c(0, 30), expand = c(0,0))

Midway_month
```
##Pearl and Hermes month graph

```{r}
PH_dat2 <- month_dat_site %>%
  filter(Site == "PHR")
PH_dat2$MonthDay <- as.Date(PH_dat2$MonthDay, format = "%m-%d")

PH_month <- ggplot(data = PH_dat2, aes(x = MonthDay, y = Mean_Turtles)) +
  geom_line() +
  scale_x_date(breaks = seq(as.Date("2023-01-01"), as.Date("2023-12-31"), by = "month"), date_labels = "%b", limits = c(as.Date("2023-01-01"), as.Date("2023-12-31")),, expand = c(0.01,0))+
  labs(x = "Month", y = "Number of Honu Seen", title = "Average Number of Honu Seen by Month at Manawai (1983-2022, n = 34)") +
  theme_minimal() +
  theme(plot.margin = margin(5.5, 10, 5.5, 5.5)) +
  annotate("rect", xmin = as.Date("2023-06-18"), xmax = as.Date("2023-07-29"), ymin = 0, ymax = 45,
           alpha = .3, fill = "orange") +
  geom_smooth(data = PH_dat2, linetype = "dashed", color = "black", se = FALSE, method = 'loess') +
  theme(panel.border = element_rect(color = "black", fill = NA)) +
  scale_y_continuous(limits = c(0, 45), breaks = seq(0, 45, by = 5), expand = c(0,0))

PH_month
```

##Kure month Graph

```{r}
Kure_dat2 <- month_dat_site %>%
  filter(Site == "KUR")
Kure_dat2$MonthDay <- as.Date(Kure_dat2$MonthDay, format = "%m-%d")

Kure_month <- ggplot(data =Kure_dat2, aes(x = MonthDay, y = Mean_Turtles)) +
  geom_line() +
  scale_x_date(breaks = seq(as.Date("2023-01-01"), as.Date("2023-12-31"), by = "month"), date_labels = "%b", limits = c(as.Date("2023-01-01"), as.Date("2023-12-31")), expand = c(0.01,0))+
  labs(x = "Month", y = "Number of Honu Seen", title = "Average Number of Honu Seen by Month at Holaniku (1983-2022, n = 32)") +
  theme_minimal() +
  annotate("rect", xmin = as.Date("2023-05-25"), xmax = as.Date("2023-07-23"), ymin = 0, ymax = 10,
           alpha = .3, fill = "yellow") +
  geom_smooth(data = Kure_dat2, linetype = "dashed", color = "black", se = FALSE, method = 'loess') +
  theme(panel.border = element_rect(color = "black", fill = NA), plot.margin = margin(5.5, 10, 5.5, 5.5)) +
  scale_y_continuous(limits = c(0, 10), breaks = seq(0, 10, by = 2), expand = c(0,0))

Kure_month
```

##Turtles seen by month - trend lines for each island

Trend line for each of the islands for the average number of turtles seen throughout the year

```{r}
month_site_trend <- ggplot() +
  geom_smooth(data = Laysan_dat2, aes(x = MonthDay, y = Mean_Turtles, color = "Kamole"), method = 'loess', se = FALSE) +
  geom_smooth(data = Lisi_dat2, aes(x = MonthDay, y = Mean_Turtles, color = "Kapou"), method = 'loess', se = FALSE) +
  geom_smooth(data = Midway_dat2, aes(x = MonthDay, y = Mean_Turtles, color = "Midway"), method = 'loess', se = FALSE) +
  geom_smooth(data = PH_dat2, aes(x = MonthDay, y = Mean_Turtles, color = "Manawai"), method = 'loess', se = FALSE) +
  geom_smooth(data = Kure_dat2, aes(x = MonthDay, y = Mean_Turtles, color = "Hōlanikū"), method = 'loess', se = FALSE) +
  labs(x = "Time", y = "Number of Turtles", title = "Mean Number of Turtles Seen Each Day") +
   scale_x_date(breaks = seq(as.Date("2023-01-01"), as.Date("2023-12-31"), by = "month"), date_labels = "%b", limits = c(as.Date("2023-01-01"), as.Date("2023-12-31")), expand = c(0.01,0)) +
  theme_minimal() +
  scale_y_continuous(limits = c(0, 25)) +
  scale_color_manual(values = c("Kamole" = "blue", "Kapou" = "red", "Manawai" = "orange", "Midway" = "green", "Hōlanikū" = "gold"), labels = c("Kamole", "Kapou", "Manawai", "Midway", "Hōlanikū"), breaks = c("Kamole", "Kapou", "Manawai", "Midway", "Hōlanikū")) +
  guides(color = guide_legend(title = "Island"))  +
  theme(legend.background = element_rect(color = "black", fill = "white"), plot.margin = margin(5.5, 5.5, 5.5, 5.5))


month_site_trend
```

##Graph of average number of samples per month for each island

I summed together the number of surveys each month in each year and then averaged each month out to find average number of samples a month. done for each island

```{r}
sample_dat <- unique_dat %>%
  mutate(Month = format(Date, "%Y-%m")) %>%
  group_by(Month, Site) %>%
  summarise(NumSurveys = n()) %>%
  mutate(Month = substr(Month, 6, 7))


avg_surveys <- sample_dat %>%
  group_by(Site, Month) %>%
  summarise(AvgSurveys = mean(NumSurveys))
avg_surveys$Month <- as.numeric(avg_surveys$Month)
custom_month_names <- c("January", "February", "March", "April", "May", "June", "July", "August", "September", "October", "November", "December")
avg_surveys$Month <- custom_month_names[avg_surveys$Month]
avg_surveys$Month <- factor(avg_surveys$Month, levels = custom_month_names, ordered = TRUE)

Kure_surveys <- avg_surveys %>%
  filter(Site == "KUR")

Laysan_surveys <- avg_surveys %>%
  filter(Site == "LAY")

Lisi_surveys <- avg_surveys %>%
  filter(Site == "LIS")

Midway_surveys <- avg_surveys %>%
  filter(Site == "MDY")

PH_surveys <- avg_surveys %>%
  filter(Site == "PHR")


# Plotting code
survey_month <- ggplot() +
  geom_line(data = Kure_surveys, aes(x = Month, y = AvgSurveys, color = "Kure", group = 1), linewidth = 0.8) +
  geom_line(data = Laysan_surveys, aes(x = Month, y = AvgSurveys, color = "Laysan", group = 1), linewidth = 0.8) +
  geom_line(data = Lisi_surveys, aes(x = Month, y = AvgSurveys, color = "Lisianski", group = 1), linewidth = 0.8) +
  geom_line(data = PH_surveys, aes(x = Month, y = AvgSurveys, color = "Pearl & Hermes", group = 1), linewidth = 0.8) +
  geom_line(data = Midway_surveys, aes(x = Month, y = AvgSurveys, color = "Midway", group = 1), linewidth = 0.8) + 
  geom_point(data = Kure_surveys, aes(x = Month, y = AvgSurveys, color = "Kure"), size = 2) +
  geom_point(data = Laysan_surveys, aes(x = Month, y = AvgSurveys, color = "Laysan"), size = 2) +
  geom_point(data = Lisi_surveys, aes(x = Month, y = AvgSurveys, color = "Lisianski"), size = 2) +
  geom_point(data = PH_surveys, aes(x = Month, y = AvgSurveys, color = "Pearl & Hermes"), size = 2) +
  geom_point(data = Midway_surveys, aes(x = Month, y = AvgSurveys, color = "Midway"), size = 2) +
  labs(x = "Time", y = "Number of Surveys", title = "Average Number of Surveys per Month") +
  scale_y_continuous(breaks = seq(0, 12, by = 2)) +
  scale_color_manual(
    name = "Island",
    values = c("Laysan" = "blue", "Lisianski" = "red", "Midway" = "green", "Pearl & Hermes" = "orange", "Kure" = "gold"),
    labels = c("Laysan", "Lisianski", "Midway", "Pearl & Hermes", "Kure"),
    breaks = c("Laysan", "Lisianski", "Midway", "Pearl & Hermes", "Kure")
  ) +
  theme_minimal() +
  theme(legend.background = element_rect(color = "black", fill = "white")) 

survey_month
```

##Total number of years where surveys happened at each island

```{r}
Laysan_years <- dat %>%
  filter(Site == "LAY")
unique_lay_years_count <- length(unique(Laysan_years$Year))

Lisi_years <- dat %>%
  filter(Site == "LIS")
unique_lis_years_count <- length(unique(Lisi_years$Year))

PH_years <- dat %>%
  filter(Site == "PHR")
unique_ph_years_count <- length(unique(PH_years$Year))

Mid_years <- dat %>%
  filter(Site == "MDY")
unique_mid_years_count <- length(unique(Mid_years$Year))

Kure_years <- dat %>%
  filter(Site == "KUR")
unique_kure_years_count <- length(unique(Kure_years$Year))
```

