---
title: "Abundance Estimates"
output: html_document
date: "2023-06-30"
---

##Libraries

```{r setup, include=FALSE}
library(readxl)
library(here)
library(dplyr)
library(tidyverse)
library(scales)
library(zoo)
```

##Data

The data I'm using to calculate abundances is all data from confirmed complete basking surveys. This is mostly coming from the HMSRP Turtle Basking Data sheet, but also from more recent sources from the other northern islands of the monument. The Lalo abundance data I'm using is from a table of estimates of NESTER abundance. However, it can be assumed that nester abundance is approximately equal to basking abundance since after nesting season, there is no basking on Lalo. So, all basking during nesting season is coming from the nesters and no year round residents. The El Nino data is coming from https://origin.cpc.ncep.noaa.gov/products/analysis_monitoring/ensostuff/ONI_v5.php 

```{r}
dat <- read_xlsx(here::here("Data", "AbundanceData.xlsx"))
Lalo_dat <- read_xlsx(here::here("Data", "LaloAbundanceData.xlsx"))
nino_dat <- read_xlsx(here::here("Data", "ONI_Data.xlsx"))
```

##Data frames for daily totals

Each data source in abundance data was different in if it reported totals of honu for each day or had a different observation for each honu. So I just made a separate data frame for each island that calculated survey sums for each day there were surveys to make it more consistent. 

```{r}
census_daily_sums <- dat %>% 
  group_by(Date, Year, Site) %>%
  summarize(Sum_Turtles = sum(Num_Turtles))

Laysan_daily_sums <- census_daily_sums %>%
  filter(Site == "LAY")

Lisi_daily_sums <- census_daily_sums %>%
  filter(Site == "LIS")

Midway_daily_sums <- census_daily_sums %>%
  filter(Site == "MDY")

PH_daily_sums <- census_daily_sums %>%
  filter(Site == "PHR")

Kure_daily_sums <- census_daily_sums %>%
  filter(Site == "KUR")

Nihoa_daily_sums <- census_daily_sums %>%
  filter(Site == "NIH")

Necker_daily_sums <- census_daily_sums %>%
  filter(Site == "NEC")
  
```

##Maximum Census Count for each year on each island

This found the maximum number of honu that were seen during a single basking survey for each year at each island. 

```{r}
Laysan_yearly_max <- Laysan_daily_sums %>%
  group_by(Year) %>%
  summarise(Yearly_Max = max(Sum_Turtles))

Lisi_yearly_max <- Lisi_daily_sums %>%
  group_by(Year) %>%
  summarise(Yearly_Max = max(Sum_Turtles))

Midway_yearly_max <- Midway_daily_sums %>%
  group_by(Year) %>%
  summarise(Yearly_Max = max(Sum_Turtles))

PH_yearly_max <- PH_daily_sums %>%
  group_by(Year) %>%
  summarise(Yearly_Max = max(Sum_Turtles))

Kure_yearly_max <- Kure_daily_sums %>%
  group_by(Year) %>%
  summarise(Yearly_Max = max(Sum_Turtles))

Nihoa_yearly_max <- Nihoa_daily_sums %>%
  group_by(Year) %>%
  summarise(Yearly_Max = max(Sum_Turtles))

Necker_yearly_max <- Necker_daily_sums %>%
  group_by(Year) %>%
  summarise(Yearly_Max = max(Sum_Turtles))
```

##Laysan Estimated Abundance Trend Line

These graphs are showing the maximum number of turtles seen in one day during one basking survey each year. It is the minimum number of basking turtles that each island has seen each year. By using the maximum number of turtles seen, we know that there is at least this many turtles, although there may be more. The code is the same for each of the northern islands.

```{r}
Laysan_yearly_max$Year <- as.numeric(Laysan_yearly_max$Year)

Laysan_est_trend <- ggplot() +
  geom_smooth(data = Laysan_yearly_max, aes(x = Year, y = Yearly_Max)) +
  geom_line(data = Laysan_yearly_max, aes (x = Year, y = Yearly_Max)) +
  labs(x= "Year", y = "Maximum Census Count", title = "Laysan Estimated Abundance (1983-2015)") +
  scale_x_continuous(name = "Year", breaks = seq(1985, 2015, by = 5), limits = c(1983, 2015), expand = c(0.01, 0)) +
  scale_y_continuous(name = "Maximum Census Count", breaks = seq(0, 18, by = 2), limits = c(0, 18)) +
  theme_minimal()

Laysan_est_trend
```

##Lisianski Estimated Abundance Trend Line

```{r}
Lisi_yearly_max$Year <- as.numeric(Lisi_yearly_max$Year)

Lisi_est_trend <- ggplot() +
  geom_smooth(data = Lisi_yearly_max, aes(x = Year, y = Yearly_Max)) +
  geom_line(data = Lisi_yearly_max, aes (x = Year, y = Yearly_Max)) +
  labs(x= "Year", y = "Maximum Census Count", title = "Lisianski Estimated Abundance (1983-2022)") + 
  scale_x_continuous(name = "Year", breaks = seq(1985, 2020, by = 5), expand = c(0, 0)) +
  scale_y_continuous(name = "Maximum Census Count", breaks = seq(0, 45, by = 5)) +
  theme_minimal()

Lisi_est_trend
```

##Midway Estimated Abundance Trend Line

```{r}
Midway_yearly_max$Year <- as.numeric(Midway_yearly_max$Year)

Midway_est_trend <- ggplot() +
  geom_smooth(data = Midway_yearly_max, aes(x = Year, y = Yearly_Max)) +
  geom_line(data = Midway_yearly_max, aes (x = Year, y = Yearly_Max)) +
  labs(x= "Year", y = "Maximum Census Count", title = "Midway Estimated Abundance (1992-2020)") +
  scale_x_continuous(name = "Year", breaks = seq(1992, 2020, by = 2), expand = c(0.01, 0)) +
  scale_y_continuous(name = "Maximum Census Count", breaks = seq(0, 35, by = 5)) +
  theme_minimal()

Midway_est_trend
```

##Pearl and Hermes Estimated Abundance Trend Line

```{r}
PH_yearly_max$Year <- as.numeric(PH_yearly_max$Year)

PH_est_trend <- ggplot() +
  geom_smooth(data = PH_yearly_max, aes(x = Year, y = Yearly_Max)) +
  geom_line(data = PH_yearly_max, aes (x = Year, y = Yearly_Max)) +
  labs(x= "Year", y = "Maximum Census Count", title = "Pearl and Hermes Estimated Abundance (1983-2022)") +
  scale_x_continuous(name = "Year", breaks = seq(1985, 2020, by = 5), expand = c(0, 0)) +
  scale_y_continuous(name = "Maximum Census Count", breaks = seq(0, 45, by = 5)) +
  theme_minimal()

PH_est_trend
```

##Kure Estimated Abundance Trend Line

```{r}
Kure_yearly_max$Year <- as.numeric(Kure_yearly_max$Year)

Kure_est_trend <- ggplot() +
  geom_smooth(data = Kure_yearly_max, aes(x = Year, y = Yearly_Max)) +
  geom_line(data = Kure_yearly_max, aes (x = Year, y = Yearly_Max)) +
  labs(x= "Year", y = "Maximum Census Count", title = "Kure Estimated Abundance (1983-2022)") +
  scale_x_continuous(name = "Year", breaks = seq(1985, 2020, by = 5), expand = c(0, 0)) +
  scale_y_continuous(name = "Maximum Census Count", breaks = seq(0, 10, by = 2), limits = c(0, 10)) +
  theme_minimal()

Kure_est_trend
```

##NWHI abundance trend

This plot is showing the smooth curves for the basking abundance data for each of the northern islands. 

```{r}
nwhi_abundance <- ggplot() +
  geom_smooth(data = Laysan_yearly_max, aes(x = Year, y = Yearly_Max, color = "Kamole/Laysan"), se = FALSE, method = 'loess') +
  geom_smooth(data = Lisi_yearly_max, aes(x = Year, y = Yearly_Max, color = "Kapou/Lisianski"), se = FALSE, method = 'loess') + 
  geom_smooth(data = Midway_yearly_max, aes(x = Year, y = Yearly_Max, color = "Midway"), se = FALSE, method = 'loess') +
  geom_smooth(data = PH_yearly_max, aes(x = Year, y = Yearly_Max, color = "Manawai/Pearl & Hermes"), se = FALSE, method = 'loess') +
  geom_smooth(data = Kure_yearly_max, aes(x = Year, y = Yearly_Max, color = "Hōlanikū/Kure"), se = FALSE, method = 'loess') +
  theme_minimal() +
  labs(x = "Time", y = "Maximum Basking Census Count", title = "Estimated Abundance of Basking Honu at Each of the NHI (1983-2022)", color = NULL) +
  scale_color_manual(values = c("Kamole/Laysan" = "blue", "Kapou/Lisianski" = "red", "Manawai/Pearl & Hermes" = "orange", "Midway" = "green", "Hōlanikū/Kure" = "gold"), labels = c("Kamole/Laysan", "Kapou/Lisianski", "Manawai/Pearl & Hermes Reef", "Midway", "Hōlanikū/Kure"), breaks = c("Kamole/Laysan", "Kapou/Lisianski", "Manawai/Pearl & Hermes", "Midway", "Hōlanikū/Kure")) +
  guides(color = guide_legend()) +
  theme(legend.background = element_rect(color = "black", fill = "white"), panel.border = element_rect(color = "black", fill = NA))+
  scale_x_continuous(limits = c(1983, 2022), breaks = seq(1984, 2022, by = 4), expand = c(0,0)) +
  scale_y_continuous(breaks = seq(0, 35, by = 5)) +
  theme(legend.position="bottom", plot.margin = margin(5.5, 10, 5.5, 5.5)) +
  guides(fill=guide_legend(nrow=1))

nwhi_abundance
```

##Total Abundance Estimate 

Here I am going to find a very rough estimate of abundance in the NHI by finding the average number of maximum baskers at each island throughout time and then summing them all together. 

```{r}
Laysan_avg <- mean(Laysan_yearly_max$Yearly_Max)
Lisi_avg <- mean(Lisi_yearly_max$Yearly_Max)
PH_avg <- mean(PH_yearly_max$Yearly_Max)
Midway_avg <- mean(Midway_yearly_max$Yearly_Max)
Kure_avg <- mean(Kure_yearly_max$Yearly_Max)

total_abundance_nhi <- sum(Laysan_avg, Lisi_avg, PH_avg, Midway_avg, Kure_avg)
total_abundance_nhi
```

##Lalo Estimated Abundance Trend Line

Fits a smooth curve and plots the data points of the Lalo estimated nester abundance data. 

```{r}
Lalo_est_trend <- ggplot() +
  geom_smooth(data = Lalo_dat, aes(x = Year, y = Estimated_Nester_Abundance)) +
  geom_line(data = Lalo_dat, aes (x = Year, y = Estimated_Nester_Abundance)) +
  labs(x= "Year", y = "Estimated Nester Abundance", title = "Lalo Estimated Nester Abundance (1973-2022)") +
  scale_x_continuous(name = "Year", breaks = seq(1974, 2022, by = 4), limits = c(1973, 2022), expand = c(0.01, 0)) +
  scale_y_continuous(name = "Estimated Nester Abundance", breaks = seq(0, 900, by = 100), limits = c(0, 900)) +
  theme_minimal()

Lalo_est_trend
```

##Abundance Estimate Trend Lines 

This graph is showing abundance estimate line graphs for each of the islands in the PMNM.

```{r}
all_islands_abundance <- ggplot() +
  geom_line(data = Laysan_yearly_max, aes(x = Year, y = Yearly_Max, color = "Laysan")) +
  geom_line(data = Lisi_yearly_max, aes(x = Year, y = Yearly_Max, color = "Lisianski")) +
  geom_line(data = Midway_yearly_max, aes(x = Year, y = Yearly_Max, color = "Midway")) +
  geom_line(data = PH_yearly_max, aes(x = Year, y = Yearly_Max, color = "Pearl & Hermes")) +
  geom_line(data = Kure_yearly_max, aes(x = Year, y = Yearly_Max, color = "Kure")) +
  geom_line(data = Lalo_dat, aes(x = Year, y = Estimated_Nester_Abundance, color = "Lalo")) +
  geom_point(data = Laysan_yearly_max, aes(x = Year, y = Yearly_Max, color = "Laysan")) +
  geom_point(data = Lisi_yearly_max, aes(x = Year, y = Yearly_Max, color = "Lisianski")) +
  geom_point(data = Midway_yearly_max, aes(x = Year, y = Yearly_Max, color = "Midway")) +
  geom_point(data = PH_yearly_max, aes(x = Year, y = Yearly_Max, color = "Pearl & Hermes")) +
  geom_point(data = Kure_yearly_max, aes(x = Year, y = Yearly_Max, color = "Kure")) +
  geom_point(data = Lalo_dat, aes(x = Year, y = Estimated_Nester_Abundance, color = "Lalo")) +
  labs(x = "Year", y = "Number of Turtles", title = "Estimated Abundance at Each Island") +
   scale_x_continuous(breaks = seq(1974,2022, by = 4), limits = c(1973, 2022), expand = c(0.01,0)) +
  theme_minimal() +
  scale_y_continuous(limits = c(0, 900), breaks = seq(0, 900, by = 50), expand = c(0,0)) +
  scale_color_manual(name = "Island", values = c("Laysan" = "blue", "Lisianski" = "red", "Midway" = "green", "Pearl & Hermes" = "orange", "Kure" = "gold", "Lalo" = "purple"), labels = c("Laysan", "Lisianski", "Midway", "Pearl & Hermes", "Kure", "Lalo"), breaks = c("Laysan", "Lisianski", "Midway", "Pearl & Hermes", "Kure", "Lalo")) +
  guides(color = guide_legend(title = "Island"))  +
  theme(legend.background = element_rect(color = "black", fill = "white"), plot.margin = margin(5.5, 5.5, 5.5, 5.5))

all_islands_abundance
```

##ONI Data

This is just me cleaning up the ONI data to where I have it in the format I want so that I can use it more easily. Then I use it to just make a line graph of the ONI index through time.  

```{r}
nino_dat_long <- nino_dat %>%
  gather(Month, ONI, -Year)

nino_dat_long$Month <- factor(nino_dat_long$Month, levels = c("JFM", "FMA", "MAM", "AMJ", "MJJ", "JJA", "JAS", "ASO", "SON", "OND", "NDJ", "DJF"))

nino_dat_sorted <- nino_dat_long %>%
  arrange(Year, Month)

month_mapping <- c("JFM" = 1, "FMA" = 2, "MAM" = 3, "AMJ" = 4, 
                   "MJJ" = 5, "JJA" = 6, "JAS" = 7, "ASO" = 8, 
                   "SON" = 9, "OND" = 10, "NDJ" = 11, "DJF" = 12)
nino_dat_sorted <- nino_dat_sorted %>%
  mutate(Month = month_mapping[Month])
nino_dat_sorted$Month <- as.numeric(nino_dat_sorted$Month)
nino_dat_sorted$YearMonth <- paste(nino_dat_sorted$Year, sprintf("%02d", nino_dat_sorted$Month), sep = "-")
nino_dat_sorted$YearMonth <- as.Date(paste(nino_dat_sorted$YearMonth, "01", sep = "-"))


oni <- ggplot(nino_dat_sorted, aes(x = YearMonth, y = ONI)) +
  geom_line() +
  theme_minimal() +
  scale_x_date(limits = c(as.Date("1950-01-01"), as.Date("2022-12-01")), expand = c(0,0))

oni
```

##Lalo plus ONI

This is where I was trying to add Lalo abundance data and ONI data to the same graph so that I could more easily tell if El Nino was having an effect on Lalo nesting abundance, but I could never quite get it to work. So this code doesn't work, but I kept it just in case. 

```{r}
Lalo_dat$Year <- as.character(Lalo_dat$Year)
Lalo_dat$Year <- as.Date(paste0(Lalo_dat$Year, "-01-01"), format = "%Y-%m-%d")

Lalo_ONI <- ggplot() +
  geom_line(data = Lalo_dat, aes (x = Year, y = Estimated_Nester_Abundance, color = "blue")) +
  geom_line(data = nino_dat_sorted, aes(x = YearMonth, y = ONI, color = "red")) +
  labs(x= "Year", y = "Estimated Nester Abundance", title = "Lalo Estimated Nester Abundance (1973-2022)") +
   scale_x_date(limits = as.Date(c("1973-01-01", "2022-12-31")), breaks = seq(as.Date("1980-01-01"), as.Date("2020-01-01"), by = "10 years"), date_labels = "%Y", expand = c(0.01, 0)) +
    scale_y_continuous(
    name = "Estimated NesterAbundance",
    breaks = seq(0, 900, by = 100),
    limits = c(0, 900),
    sec.axis = sec_axis(
      trans = ~ rescale(., to = c(-2, 2)),
      name = "ONI",
      breaks = c(-2, -1, 0, 1, 2),
      labels = c("-2", "-1", "0", "1", "2")
    )) +
  theme_minimal() +
  scale_color_manual(values = c("blue", "red"), labels = c("Lalo", "ONI"), name = "thing")
  

Lalo_ONI
```









