---
title: "Size_Island_Year_Sex"
output: html_document
date: "2023-06-16"
---

## Libraries

```{r}
library(here)
library(tidyverse)
library(readxl)
library(lme4)
library(dplyr)
```

## Data

The data I used for this analysis is coming from the MTBAP database and the more recent 2022 data that hasn't been entered into the database yet. I decided to use CCL to compare sizes of honu between islands, since that is the metric that was most often collected. I was using only adult honu, which I defined as above 81 cm in SCL which is coming from an adult classification metric from Balazs 1980. Sometimes, honu in the database only had one or two measurements. I used an equation that is usually used for Atlantic greens to estimate SCL based on CCL and vice versa. The equation is CCL = (SCL + 0.0515)/0.9426 (Goshe et al. 2010). So as long as a turt had either SCL or CCL, it could be included in my analysis. 

```{r setup, include=FALSE}

dat <- SizeData
head(dat)

```

## Table of Averages

I found the average SCL, CCL, SCW, and CCW for honu at each island for each year separated out by males and females. 

```{r}
##stores all size metrics as numbers (instead of characters lol)
dat$SCL <- as.numeric(dat$SCL)
dat$CCL <- as.numeric(dat$CCL)
dat$SCW <- as.numeric(dat$SCW)
dat$CCW <- as.numeric(dat$CCW)
dat$Year <- as.numeric(dat$Year)


##makes four different tables storing the averages for each of the four size metrics for each year for each island

SCL_avgs <- dat %>%
  filter(!is.na(SCL)) %>%
  group_by(Island, Sex, Year) %>%
  dplyr::summarise(SCL.avg = mean(SCL, na.rm = TRUE), sample_size = n())
  
CCL_avgs <- dat %>%
  group_by(Island, Sex, Year) %>%
  dplyr::summarise(CCL.avg = mean(c(CCL, Est_CCL), na.rm = TRUE), sample_size = sum(!is.na(CCL)|!is.na(Est_CCL)))

SCW_avgs <- dat %>%
  filter(!is.na(SCW)) %>%
  group_by(Island, Sex, Year) %>%
  dplyr::summarise(SCW.avg = mean(SCW, na.rm = TRUE), sample_size = n())

CCW_avgs <- dat %>%
  filter(!is.na(CCW)) %>%
  group_by(Island, Sex, Year) %>%
  dplyr::summarise(CCW.avg = mean(CCW, na.rm = TRUE), sample_size = n())

```
## Laysan bar graph CCL averages

Then I made bar graphs for the average CCLs for Laysan separated by males and females. I did the same thing for each island. 

```{r}
##new table for just the Laysan data I want to use in the graph with certain years from the CCL avgs table
Laysan_dat <- CCL_avgs %>%
  filter(Island == "LAYS") %>%
  filter(Year %in% c(1968, 1982, 1984, 1993))

Laysan_dat$Year <- as.character(Laysan_dat$Year)

##graph for Laysan averages of CCL 
Laysan <- ggplot(Laysan_dat, aes(x = Year, y = CCL.avg, fill = Sex)) +
  geom_col(position = position_dodge(), width = 0.5) +
  labs(x = "Year", y = "Average CCL (cm)", title = "Laysan CCL averages") +
  theme_minimal() +
  scale_fill_manual(values = c("#F5495E", "#539DEE"), labels = c("Female", "Male")) +
  geom_text(aes(label = round(CCL.avg, digits = 1)), fontface = "bold", vjust = -0.3, position = position_dodge(.5), size = 4) +
  geom_text(aes(label = paste0("n = ", sample_size)), vjust = 1.5, position = position_dodge(.5), size = 3) +
  scale_y_continuous(minor_breaks = seq(0, 100, 5)) +
  theme(panel.grid.major = element_line(size = 1, color = "gray90")) +
  theme(legend.spacing.y = unit(2, "mm"), axis.text = element_text(colour = 1, size = 12),
        legend.background = element_blank(),
        legend.box.background = element_rect(colour = "black"))
```

##Lisianski bar graph CCL averages

```{r}
##new table for just the Lisianski data I want to use in the graph with certain years from the CCL avgs table
Lisi_dat <- CCL_avgs %>%
  filter(Island == "LISI") %>%
  filter(Year %in% c(1967, 1968, 1982, 1991, 1992, 2018, 2022))

Lisi_dat$Year <- as.character(Lisi_dat$Year)

##graph for Lisi averages of CCL 
Lisi <- ggplot(Lisi_dat, aes(x = Year, y = CCL.avg, fill = Sex)) +
  geom_col(position = position_dodge(), width = 0.7) +
  labs(x = "Year", y = "Average CCL (cm)", title = "Lisi CCL averages") +
  theme_minimal() +
  scale_fill_manual(values = c("#F5495E", "#539DEE"), labels = c("Female", "Male")) +
  geom_text(aes(label = round(CCL.avg, digits = 1)), fontface = "bold", vjust = -0.3, position = position_dodge(.7), size = 4) +
  geom_text(aes(label = paste0("n = ", sample_size)), vjust = 1.5, position = position_dodge(.7), size = 3) +
  scale_y_continuous(minor_breaks = seq(0, 100, 5)) +
  theme(panel.grid.major = element_line(size = 1, color = "gray90")) +
  theme(legend.spacing.y = unit(2, "mm"), axis.text = element_text(colour = 1, size = 12),
        legend.background = element_blank(),
        legend.box.background = element_rect(colour = "black"))

```

##Midway Bar Graph CCL averages

```{r}
##new table for just the Midway data I want to use in the graph with certain years from the CCL avgs table
Midway_dat <- CCL_avgs %>%
  filter(Island == "MIDW") %>%
  filter(Year %in% c(1998, 1999, 2000, 2001, 2019, 2020, 2022))

Midway_dat$Year <- as.character(Midway_dat$Year)

##graph for Midway averages of CCL 
Midway <- ggplot(Midway_dat, aes(x = Year, y = CCL.avg, fill = Sex)) +
  geom_col(position = position_dodge(), width = 0.75) +
  labs(x = "Year", y = "Average CCL (cm)", title = "Midway CCL averages") +
  theme_minimal() +
  scale_fill_manual(values = c("#F5495E", "#539DEE"), labels = c("Female", "Male")) +
  geom_text(aes(label = round(CCL.avg, digits = 1)), fontface = "bold", vjust = -0.3, position = position_dodge(0.75), size = 4) +
  geom_text(aes(label = paste0("n = ", sample_size)), vjust = 1.5, position = position_dodge(0.75), size = 3) +
  scale_y_continuous(minor_breaks = seq(0, 100, 5)) +
  theme(panel.grid.major = element_line(size = 1, color = "gray90")) +
  theme(legend.spacing.y = unit(2, "mm"), axis.text = element_text(colour = 1, size = 12),
        legend.background = element_blank(),
        legend.box.background = element_rect(colour = "black"))

```

##Pearl and Hermes Bar Graph CCL averages

```{r}
##new table for just the Pearl and Hermes data I want to use in the graph with certain years from the CCL avgs table
PH_dat <- CCL_avgs %>%
  filter(Island == "PEAR") %>%
  filter(Year %in% c(1966, 1967, 1968, 1970, 1982, 1983, 1984, 1986, 1988, 1991, 2022))

PH_dat$Year <- as.character(PH_dat$Year)

##graph for Laysan averages of CCL 
PH <- ggplot(PH_dat, aes(x = Year, y = CCL.avg, fill = Sex)) +
  geom_col(position = position_dodge(), width = 0.8) +
  labs(x = "Year", y = "Average CCL (cm)", title = "Pearl & Hermes CCL averages") +
  theme_minimal() +
  scale_fill_manual(values = c("#F5495E", "#539DEE"), labels = c("Female", "Male")) +
  geom_text(aes(label = round(CCL.avg, digits = 1)), fontface = "bold", vjust = -0.3, position = position_dodge(.8), size = 4) +
  geom_text(aes(label = paste0("n = ", sample_size)), vjust = 1.5, position = position_dodge(.8), size = 3) +
  scale_y_continuous(minor_breaks = seq(0, 100, 5)) +
  theme(panel.grid.major = element_line(size = 1, color = "gray90")) +
  theme(legend.spacing.y = unit(2, "mm"), axis.text = element_text(colour = 1, size = 12),
        legend.background = element_blank(),
        legend.box.background = element_rect(colour = "black"), aspect.ratio = 5/9) 
```


##Laysan Trend Line CCL Averages

Then I realized that trend lines would actually probably be more helpful and easier to look at, so then I made trend lines instead for each of the islands. 

```{r}
##Separating out males and females 
Laysan_trend_dat_F <- CCL_avgs %>%
  filter(Island == "LAYS", Sex == "F")
Laysan_trend_dat_M <- CCL_avgs %>%
  filter(Island == "LAYS", Sex == "M")

##Graph with data line and trend line
Laysan_trend <- ggplot() +
  geom_smooth(data = Laysan_trend_dat_F, aes(x = Year, y = CCL.avg), color = "chocolate1", linetype = "dashed", method = "lm", se = FALSE, formula = y ~ x) +
  geom_smooth(data = Laysan_trend_dat_M, aes(x = Year, y = CCL.avg), color = "blue", linetype = "dashed", method = "lm", se = FALSE, formula = y ~ x) +
  geom_line(data = Laysan_trend_dat_F, aes(x = Year, y = CCL.avg, color = "Female"), linewidth = 1) +
  geom_line(data = Laysan_trend_dat_M, aes(x = Year, y = CCL.avg, color = "Male"), linewidth = 1) +
  scale_x_continuous(name = "Year", breaks = seq(1970, 2020, by = 10), limits = c(1966, 2022), minor_breaks = seq(1966, 2022, by = 2), expand = c(0, 0)) +
  scale_y_continuous(name = "Average CCL (cm)", breaks = seq(86, 104, by = 2), limits = c(85, 105)) +
  labs(title = "Kamole Average CCL (cm) Trend of Adult Honu 1966 - 2022", size = "Sample Size") +
  theme_minimal() +
  theme(
    axis.text.x = element_text(face = "bold"),
    axis.text.y = element_text(face = "bold"),
    panel.grid.major = element_line(size = 1, color = "gray90"),
    legend.background = element_rect(color = "black", fill = "white"),
    legend.position = "right"  # Set legend position to the right side
  ) +
  scale_color_manual(
    name = "Sex",
    values = c("Female" = "chocolate1", "Male" = "blue"),
    labels = c("Female", "Male")
  ) +
  guides(
    color = guide_legend(override.aes = list(size = 4, shape = NA)),
    size = guide_legend(override.aes = list(color = "black"))
  ) +
  coord_cartesian(clip = "off") +
  theme(panel.border = element_rect(color = "black", fill = NA)) + 
  geom_point(data = Laysan_trend_dat_F, aes(x = Year, y = CCL.avg, size = sample_size, color = "Female"), show.legend = TRUE) +
  geom_point(data = Laysan_trend_dat_M, aes(x = Year, y = CCL.avg, size = sample_size, color = "Male"), show.legend = TRUE) 


```

##Lisianski Trend Line CCL Averages

```{r}
Lisi_trend_dat_F <- CCL_avgs %>%
  filter(Island == "LISI", Sex == "F")
Lisi_trend_dat_M <- CCL_avgs %>%
  filter(Island == "LISI", Sex == "M")

Lisi_trend <- ggplot() +
  geom_smooth(data = Lisi_trend_dat_F, aes(x = Year, y = CCL.avg), color = "chocolate1", linetype = "dashed", method = "lm", se = FALSE, formula = y ~ x) +
  geom_smooth(data = Lisi_trend_dat_M, aes(x = Year, y = CCL.avg), color = "blue", linetype = "dashed", method = "lm", se = FALSE, formula = y ~ x) +
  geom_line(data = Lisi_trend_dat_F, aes(x = Year, y = CCL.avg, color = "Female"), linewidth = 1) +
  geom_line(data = Lisi_trend_dat_M, aes(x = Year, y = CCL.avg, color = "Male"), linewidth = 1) +
  scale_x_continuous(name = "Year", breaks = seq(1970, 2020, by = 10), limits = c(1967, 2022), minor_breaks = seq(1968, 2022, by = 2), expand = c(0, 0)) +
  scale_y_continuous(name = "Average CCL (cm)", breaks = seq(86, 104, by = 2), limits = c(85, 105)) +
  labs(title = "Kapou Average CCL (cm) Trend of Adult Honu 1967 - 2022", size = "Sample Size") +
  theme_minimal() +
  theme(
    axis.text.x = element_text(face = "bold"),
    axis.text.y = element_text(face = "bold"),
    panel.grid.major = element_line(size = 1, color = "gray90"),
    legend.background = element_rect(color = "black", fill = "white"),
    legend.position = "right"  # Set legend position to the right side
  ) +
  scale_color_manual(
    name = "Sex",
    values = c("Female" = "chocolate1", "Male" = "blue"),
    labels = c("Female", "Male")
  ) +
  guides(
    color = guide_legend(override.aes = list(size = 4, shape = NA)),
    size = guide_legend(override.aes = list(color = "black"))
  ) +
  coord_cartesian(clip = "off") +
  theme(panel.border = element_rect(color = "black", fill = NA)) +
  geom_point(data = Lisi_trend_dat_F, aes(x = Year, y = CCL.avg, size = sample_size, color = "Female"), show.legend = TRUE) +
  geom_point(data = Lisi_trend_dat_M, aes(x = Year, y = CCL.avg, size = sample_size, color = "Male"), show.legend = TRUE) 

```

##Midway Trend Line CCL Averages

```{r}
Midway_trend_dat_F <- CCL_avgs %>%
  filter(Island == "MIDW", Sex == "F")
Midway_trend_dat_M <- CCL_avgs %>%
  filter(Island == "MIDW", Sex == "M")

Midway_trend <- ggplot() +
  geom_smooth(data = Midway_trend_dat_F, aes(x = Year, y = CCL.avg), color = "chocolate1", linetype = "dashed", method = "lm", se = FALSE, formula = y ~ x) +
  geom_smooth(data = Midway_trend_dat_M, aes(x = Year, y = CCL.avg), color = "blue", linetype = "dashed", method = "lm", se = FALSE, formula = y ~ x) +
  geom_line(data = Midway_trend_dat_F, aes(x = Year, y = CCL.avg, color = "Female"), linewidth = 1) +
  geom_line(data = Midway_trend_dat_M, aes(x = Year, y = CCL.avg, color = "Male"), linewidth = 1) +
  scale_x_continuous(name = "Year", breaks = seq(2000, 2020, by = 4), limits = c(1997, 2022), minor_breaks = seq(1997, 2022, by = 1), expand = c(0, 0)) +
  scale_y_continuous(name = "Average CCL (cm)", breaks = seq(86, 104, by = 2), limits = c(85, 105)) +
  labs(title = "Midway Average CCL (cm) Trend of Adult Honu 1997 - 2022", size = "Sample Size") +
  theme_minimal() +
  theme(
    axis.text.x = element_text(face = "bold"),
    axis.text.y = element_text(face = "bold"),
    panel.grid.major = element_line(size = 1, color = "gray90"),
    legend.background = element_rect(color = "black", fill = "white"),
    legend.position = "right"  # Set legend position to the right side
  ) +
  scale_color_manual(
    name = "Sex",
    values = c("Female" = "chocolate1", "Male" = "blue"),
    labels = c("Female", "Male")
  ) +
  guides(
    color = guide_legend(override.aes = list(size = 4, shape = NA)),
    size = guide_legend(override.aes = list(color = "black"))
  ) +
  coord_cartesian(clip = "off") +
  theme(panel.border = element_rect(color = "black", fill = NA)) +
  geom_point(data = Midway_trend_dat_F, aes(x = Year, y = CCL.avg, size = sample_size, color = "Female"), show.legend = TRUE) +
  geom_point(data = Midway_trend_dat_M, aes(x = Year, y = CCL.avg, size = sample_size, color = "Male"), show.legend = TRUE) 

```

##Pearl and Hermes Trend Line CCL Averages

```{r}
PH_trend_dat_F <- CCL_avgs %>%
  filter(Island == "PEAR", Sex == "F")
PH_trend_dat_F_clean <- na.omit(PH_trend_dat_F)
PH_trend_dat_M <- CCL_avgs %>%
  filter(Island == "PEAR", Sex == "M")

PH_trend <- ggplot() +
  geom_smooth(data = PH_trend_dat_F, aes(x = Year, y = CCL.avg), color = "chocolate1", linetype = "dashed", method = "lm", se = FALSE, formula = y ~ x) +
  geom_smooth(data = PH_trend_dat_M, aes(x = Year, y = CCL.avg), color = "blue", linetype = "dashed", method = "lm", se = FALSE, formula = y ~ x) +
  geom_line(data = PH_trend_dat_F, aes(x = Year, y = CCL.avg, color = "Female"), linewidth = 1) +
  geom_line(data = PH_trend_dat_M, aes(x = Year, y = CCL.avg, color = "Male"), linewidth = 1) +
  scale_x_continuous(name = "Year", breaks = seq(1970, 2020, by = 10), limits = c(1964, 2022), minor_breaks = seq(1964, 2022, by = 2), expand = c(0, 0)) +
  scale_y_continuous(name = "Average CCL (cm)", breaks = seq(86, 104, by = 2), limits = c(85, 105)) +
  labs(title = "Manawai CCL (cm) Trend of Adult Honu 1964 - 2022", size = "Sample Size") +
  theme_minimal() +
  theme(
    axis.text.x = element_text(face = "bold"),
    axis.text.y = element_text(face = "bold"),
    panel.grid.major = element_line(size = 1, color = "gray90"),
    legend.background = element_rect(color = "black", fill = "white"),
    legend.position = "right"  # Set legend position to the right side
  ) +
  scale_color_manual(
    name = "Sex",
    values = c("Female" = "chocolate1", "Male" = "blue"),
    labels = c("Female", "Male")
  ) +
  guides(
    color = guide_legend(override.aes = list(size = 4, shape = NA)),
    size = guide_legend(override.aes = list(color = "black"))
  ) +
  coord_cartesian(clip = "off") +
  theme(panel.border = element_rect(color = "black", fill = NA)) +
  geom_point(data = PH_trend_dat_F, aes(x = Year, y = CCL.avg, size = sample_size, color = "Female"), show.legend = TRUE) +
  geom_point(data = PH_trend_dat_M, aes(x = Year, y = CCL.avg, size = sample_size, color = "Male"), show.legend = TRUE) 

```

##Table of all sample sizes

Then I found all the total sample sizes of males and females at each island and put them in a table for easy reference. They are not separated out by year, but the bar graphs have specific sample sizes listed out for each year and then trend lines have relative sample size shown using differently sized points on the line graph.

```{r}
Islands <- c(" Total Laysan", "Total Lisianski", "Total Midway", "Total Pearl & Hermes", "Laysan Males", "Laysan Females", "Lisianski Males", "Lisianski Females", "Midway Males", "Midway Females", "Pearl & Hermes Males", "Pearl & Hermes Females")
Samples<- c((sum(Laysan_trend_dat_F$sample_size) +sum(Laysan_trend_dat_M$sample_size)), (sum(Lisi_trend_dat_F$sample_size) + sum(Lisi_trend_dat_M$sample_size)), (sum(Midway_trend_dat_F$sample_size) + sum(Midway_trend_dat_M$sample_size)), (sum(PH_trend_dat_F$sample_size) + sum(PH_trend_dat_M$sample_size)), sum(Laysan_trend_dat_M$sample_size), sum(Laysan_trend_dat_F$sample_size), sum(Lisi_trend_dat_M$sample_size), sum(Lisi_trend_dat_F$sample_size), sum(Midway_trend_dat_M$sample_size), sum(Midway_trend_dat_F$sample_size), sum(PH_trend_dat_M$sample_size), sum(PH_trend_dat_F$sample_size))
totals <- data.frame(Islands, Samples)
```

