---
title: "Lalo_CCL_trends"
output: html_document
date: "2023-06-22"
---

##Libraries

```{r setup, include=FALSE}
library(here)
library(tidyverse)
library(readxl)
library(lme4)
library(dplyr)
```

##Data

Here, I'm using the same data that I did for the individual NHI CCL trends, but I also have another data set with all of the size data from Lalo. I cleaned it using the same metrics as the other data set (so only adults > 81 cm SCL and those with sex information). 

```{r}
dat <- SizeData
Lalo_dat <- combined_lalo
head(Lalo_dat)
```

##Remove Duplicates

The Lalo data had a lot of duplicates for each turtle since we see each turtle many times in one year, so I removed the duplicates of each Turtle ID if it was seen during the same year. That way, each turtle is only being counted in the math once. 

```{r}
Lalo_dat <- Lalo_dat[!duplicated(Lalo_dat[c("Turtle ID", "Year")]), ]
```


## Table of Averages

Found the average CCL of Lalo male and female honu for each year.

```{r}
Lalo_CCL_avgs <- Lalo_dat %>%
  group_by(Sex, Year) %>%
  dplyr::summarise(CCL.avg = mean(c(CCL, Est_CCL), na.rm = TRUE), sample_size = sum(!is.na(CCL)|!is.na(Est_CCL)))

```

##Lalo CCL Averages Bar Graph

Bar graph for Lalo's CCL averages. 

```{r}
Lalo_CCL_avgs$Year <- as.character(Lalo_CCL_avgs$Year)

Lalo_bar <- Lalo_CCL_avgs %>%
  filter(Year %in% c(1965, 1966, 1967, 1968, 1975, 1994, 1999, 2016, 2017, 2018, 2019, 2021, 2022))

##graph for Lalo averages of CCL 
Lalo <- ggplot(Lalo_bar, aes(x = Year, y = CCL.avg, fill = Sex)) +
  geom_col(position = position_dodge(), width = 0.8) +
  labs(x = "Year", y = "Average CCL (cm)", title = "Lalo CCL averages") +
  theme_minimal() +
  scale_fill_manual(values = c("#F5495E", "#539DEE"), labels = c("Female", "Male")) +
  geom_text(aes(label = round(CCL.avg, digits = 1)), fontface = "bold", vjust = -0.3, position = position_dodge(.8), size = 4) +
  geom_text(aes(label = paste0("n = ", sample_size)), vjust = 1.5, position = position_dodge(.8), size = 3) +
  scale_y_continuous(minor_breaks = seq(0, 100, 5)) +
  theme(panel.grid.major = element_line(size = 1, color = "gray90")) +
  theme(legend.spacing.y = unit(2, "mm"), axis.text = element_text(colour = 1, size = 12),
        legend.background = element_blank(),
        legend.box.background = element_rect(colour = "black"))
```

##Lalo Trend Line CCL averages

Trend line for Lalo's CCL averages. 

```{r}
Lalo_CCL_avgs$Year <- as.numeric(Lalo_CCL_avgs$Year)

##Separating out males and females 
Lalo_trend_dat_F <- Lalo_CCL_avgs %>%
  filter(Sex == "F")
Lalo_trend_dat_M <- Lalo_CCL_avgs %>%
  filter(Sex == "M")

##Graph with data line and trend line
Lalo_trend <- ggplot() +
  geom_smooth(data = Lalo_trend_dat_F, aes(x = Year, y = CCL.avg), color = "red", linetype = "dashed", method = "lm", se = FALSE, formula = y ~ x) +
  geom_smooth(data = Lalo_trend_dat_M, aes(x = Year, y = CCL.avg), color = "blue", linetype = "dashed", method = "lm", se = FALSE, formula = y ~ x) +
  geom_line(data = Lalo_trend_dat_F, aes(x = Year, y = CCL.avg, color = "Female"), linewidth = 1) +
  geom_line(data = Lalo_trend_dat_M, aes(x = Year, y = CCL.avg, color = "Male"), linewidth = 1) +
  scale_x_continuous(name = "Year", breaks = seq(1970, 2020, by = 10), limits = c(1965, 2022), minor_breaks = seq(1965, 2022, by = 5), expand = c(0, 0)) +
  scale_y_continuous(name = "Average CCL (cm)", breaks = seq(88, 104, by = 2), limits = c(88, 105)) +
  labs(title = "Lalo Average CCL (cm) Trend 1965 - 2022", size = "Sample Size") +
  theme_minimal() +
  theme(
    axis.text.x = element_text(face = "bold"),
    axis.text.y = element_text(face = "bold"),
    panel.grid.major = element_line(size = 1, color = "gray90"),
    legend.background = element_rect(color = "black", fill = "white"),
    legend.position = "right"  # Set legend position to the right side
  ) +
  scale_color_manual(
    name = "Sex",
    values = c("Female" = "red", "Male" = "blue"),
    labels = c("Female", "Male")
  ) +
  guides(
    color = guide_legend(override.aes = list(size = 4, shape = NA)),
    size = guide_legend(override.aes = list(color = "black"))
  ) +
  coord_cartesian(clip = "off") +
  geom_point(data = Lalo_trend_dat_F, aes(x = Year, y = CCL.avg, size = sample_size, color = "Female"), show.legend = TRUE) +
  geom_point(data = Lalo_trend_dat_M, aes(x = Year, y = CCL.avg, size = sample_size, color = "Male"), show.legend = TRUE)
```


##Lalo vs all other NHI trend line (keeping the sexes distinct)

This graph is to compare all of the other northern Hawaiian islands to just Lalo to see if there is a difference between the size of honu at Lalo versus the rest of the monument. In this graph, I kept the sexes distinct to see if there was a difference in size between the males and females of each region (spoiler alert: there's a significant difference in females but not in the males).

```{r}
#New dataframe of table of averages of CCL just by year so combining all the islands 
NWHI_CCL_avgs <- dat %>%
  group_by(Sex, Year) %>%
  dplyr::summarise(CCL.avg = mean(c(CCL, Est_CCL), na.rm = TRUE), sample_size = sum(!is.na(CCL)|!is.na(Est_CCL)))

NWHI_CCL_avgs$Year <- as.numeric(NWHI_CCL_avgs$Year)

NWHI_CCL_avgs_F <- NWHI_CCL_avgs %>%
  filter(Sex == "F")

NWHI_CCL_avgs_M <- NWHI_CCL_avgs %>%
  filter(Sex == "M")

##Trend line of NWHI and Lalo 
Lalo_NWHI_trend_w_sex <- ggplot() +
  geom_line(data = Lalo_trend_dat_F, aes(x = Year, y = CCL.avg, color = "Laysan Female"), linewidth = 1) +
  geom_line(data = Lalo_trend_dat_M, aes(x = Year, y = CCL.avg, color = "Laysan Male"), linewidth = 1) +
  geom_line(data = NWHI_CCL_avgs_F, aes(x = Year, y = CCL.avg, color = "NHI Females"), linewidth = 1) +
  geom_line(data = NWHI_CCL_avgs_M, aes(x = Year, y = CCL.avg, color = "NHI Males"), linewidth = 1) +
  geom_smooth(data = Lalo_trend_dat_F, aes(x = Year, y = CCL.avg), color = "deeppink1", linetype = "dashed", method = "lm", se = FALSE, formula = y ~ x) +
  geom_smooth(data = Lalo_trend_dat_M, aes(x = Year, y = CCL.avg), color = "plum1", linetype = "dashed", method = "lm", se = FALSE, formula = y ~ x) +
  geom_smooth(data = NWHI_CCL_avgs_F, aes(x = Year, y = CCL.avg), color = "black", linetype = "dashed", method = "lm", se = FALSE, formula = y ~ x) +
  geom_smooth(data = NWHI_CCL_avgs_M, aes(x = Year, y = CCL.avg), color = "#BFC9CA", linetype = "dashed", method = "lm", se = FALSE, formula = y ~ x) +
  scale_x_continuous(name = "Year", breaks = seq(1970, 2020, by = 10), limits = c(1966, 2022), minor_breaks = seq(1966, 2022, by = 2), expand = c(0, 0)) +
  scale_y_continuous(name = "Average CCL (cm)", breaks = seq(86, 104, by = 2), limits = c(85, 105)) +
  labs(title = "Lalo vs NHI Average CCL (cm) Trends of Adult Honu (1966 - 2022)") +
  theme_minimal() +
  theme(
    axis.text.x = element_text(face = "bold"),
    axis.text.y = element_text(face = "bold"),
    panel.grid.major = element_line(size = 1, color = "gray90"),
    legend.background = element_rect(color = "black", fill = "white"),
    legend.box.background = element_rect(color = "black", fill = "white"),
    legend.position = "right"  # Set legend position to the right side
  ) +
  scale_color_manual(
    name = "Region and Sex",
    values = c("deeppink1", "plum1", "black", "#BFC9CA"),
    labels = c("Lalo Females", "Lalo Males", "NHI Females", "NHI Males")
  ) +
  geom_point(data = Lalo_trend_dat_F, aes(x = Year, y = CCL.avg), color = "deeppink1", size = 1.5) +
  geom_point(data = Lalo_trend_dat_M, aes(x = Year, y = CCL.avg), color = "plum1", size = 1.5) +
  geom_point(data = NWHI_CCL_avgs_F, aes(x = Year, y = CCL.avg), color = "black", size = 1.5) +
  geom_point(data = NWHI_CCL_avgs_M, aes(x = Year, y = CCL.avg), color = "#BFC9CA", size = 1.5) +
  theme(panel.border = element_rect(color = "black", fill = NA)) +
  theme(legend.position="bottom", plot.margin = margin(5.5, 10, 5.5, 5.5)) +
  guides(fill=guide_legend(nrow=1))
  

```
##Lalo vs NWHI Trend Line (combining sexes here for simplicity)

Here, I combined the sexes just to make it easier to look at. However, since there's a significant different between the NHI and Lalo females and not the males, then it's probably only worth looking at the above graph to get that specificity. 

```{r}
NWHI_trend2 <- dat %>%
  group_by(Year) %>%
  dplyr::summarise(CCL.avg = mean(c(as.numeric(CCL), as.numeric(Est_CCL)), na.rm = TRUE), sample_size = sum(!is.na(CCL) | !is.na(Est_CCL)))

Lalo_trend2 <- Lalo_dat %>%
  group_by(Year) %>%
  dplyr::summarise(CCL.avg = mean(c(as.numeric(CCL), as.numeric(Est_CCL)), na.rm = TRUE), sample_size = sum(!is.na(CCL) | !is.na(Est_CCL)))

Lalo_trend2$Year <- as.numeric(Lalo_trend2$Year)
NWHI_trend2$Year <- as.numeric(NWHI_trend2$Year)

##Trend Line for NWHI and Lalo CCL averages with combined sexes
Lalo_NWHI_trend_combined <- ggplot() +
  geom_line(data = Lalo_trend2, aes(x = Year, y = CCL.avg, color = "Lalo"), linewidth = 1) +
  geom_line(data = NWHI_trend2, aes(x = Year, y = CCL.avg, color = "NWHI"), linewidth = 1) +
  geom_smooth(data = Lalo_trend2, aes(x = Year, y = CCL.avg), color = "orange", linetype = "dashed", method = "lm", se = FALSE, formula = y ~ x) +
  geom_smooth(data = NWHI_trend2, aes(x = Year, y = CCL.avg), color = "limegreen", linetype = "dashed", method = "lm", se = FALSE, formula = y ~ x) +
  scale_x_continuous(name = "Year", breaks = seq(1970, 2020, by = 10), limits = c(1966, 2022), minor_breaks = seq(1966, 2022, by = 2), expand = c(0, 0)) +
  scale_y_continuous(name = "Average CCL (cm)", breaks = seq(86, 104, by = 2), limits = c(85, 105)) +
  labs(title = "Lalo vs NWHI Average CCL (cm) Trends 1966 - 2022") +
  theme_minimal() +
  theme(
    axis.text.x = element_text(face = "bold"),
    axis.text.y = element_text(face = "bold"),
    panel.grid.major = element_line(size = 1, color = "gray90"),
    legend.background = element_rect(color = "black", fill = "white"),
    legend.box.background = element_rect(color = "black", fill = "white"),
    legend.position = "right"  # Set legend position to the right side
  ) +
  scale_color_manual(
    name = "Region",
    values = c("orange", "limegreen"),
    labels = c("Lalo", "NWHI")
  ) +
  geom_point(data = Lalo_trend2, aes(x = Year, y = CCL.avg), color = "orange", size = 1.5) +
  geom_point(data = NWHI_trend2, aes(x = Year, y = CCL.avg), color = "limegreen", size = 1.5)

```

##Table of Sample Sizes

Again, just made a table of total sample sizes for the NHI and Lalo males and females for easy reference. The yearly sample sizes for Lalo are specifically listed in the bar graph and then in the trend line, there are various sized dots to show relative sample size. 

```{r}
Region <- c("Lalo", "All NWHI Islands", "Lalo Males", "Lalo Females", "All NWHI Males", "All NWHI Females")
Samples <- c(sum(Lalo_CCL_avgs$sample_size), sum(NWHI_CCL_avgs$sample_size), sum(Lalo_trend_dat_M$sample_size), sum(Lalo_trend_dat_F$sample_size), sum(NWHI_CCL_avgs_M$sample_size), sum(NWHI_CCL_avgs_F$sample_size))
Lalo_totals <- data.frame(Region, Samples)
```

