---
title: "Resights"
output: html_document
date: "2023-07-10"
---

##Libraries

```{r setup, include=FALSE}
library(here)
library(dplyr)
library(tidyverse)
library(readxl)
```

##Data

The data I'm using for this part of my project is coming from the MTBAP database and the 2022 and 2021 data sources that haven't yet been entered.

```{r}
event_dat <- read_xlsx(here::here("Data", "combined_database2.xlsx"))

event_dat <- event_dat %>%
  filter(!is.na(Turtle_ID))

event_dat$Year <- as.numeric(event_dat$Year)

```

##Finding Unique Turtle IDs in the database and storing them together in their own dataframe

This code creates a new data frame for each unique turtle id and puts all the observations/events with that turtle id into the same data frame so that you can look at all the sightings of one turtle at one time. This includes turtles that have just one observation.

```{r}
#creates new list to store all the new dataframes for each individual turtle
duplicate_ids <- list()

#gets unique turtle id values
unique_turtle_ids <- unique(event_dat$Turtle_ID)

#for loop to iterate over each unique turtle id from unique_turtle_ids
for (id in unique_turtle_ids) {
  #puts each unique turtle id in a new row
  duplicate_rows <- event_dat[event_dat$Turtle_ID == id, ]
  #if statement checks for duplicates 
  if (nrow(duplicate_rows) >= 1) {
    #stores the duplicate rows in a dataframe in the duplicate_ids list
    duplicate_ids[[id]] <- duplicate_rows
  }
}

##You can put any turtle ID in the quotations and get every instance of it in the database
specific_id <- duplicate_ids[['9012']]

specific_id

```

##Finding turtles that moved 

Now I want to only look at turtles that moved between islands. Turtles that stayed at the same island are also interesting, but not what I'm looking at right now. You can also find a specific turtle that moved with the specific_df line. 140 total honu moved between islands. 

```{r}

moving_turtles <- list()

# Iterate over each dataframe in the duplicate_dfs list (df_name in names(duplicate_ids) line obtains each name of the dataframes in duplicate_ids using the names() function)
for (df_name in names(duplicate_ids)) {
  # Get the current dataframe
  current_df <- duplicate_ids[[df_name]]
  # Check if there are multiple unique values in the Islands column
  unique_islands <- unique(current_df$Island)
  if (length(unique_islands) > 1) {
    # Store the dataframe in the moving_turtles list
    moving_turtles[[df_name]] <- current_df
  }
}

specific_df <- moving_turtles[['411C73127B']]

specific_df
```

##Turtles with dual citizenship

Just from scrolling through the data frames, it seems like of the turtles that moved, most of them only moved between two islands. I want to know how many turtles go between only two islands. (Spoiler alert: it was 139 out of the 140 moving turtles). Again you can quickly pull up the data frame for turtle with dual citizenship using specific_cit. 

```{r}
dual_citizen <- list()

for(df_name in names(duplicate_ids)) {
  current_df <- duplicate_ids[[df_name]]
  unique_islands <- unique(current_df$Island)
  if(length(unique_islands) == 2) {
    dual_citizen[[df_name]] <- current_df
  }
}

specific_cit <- dual_citizen[['1727']]
specific_cit

```

##Triple citizen???

Only one turtle in the entire history of the program has been seen at more than two islands. She was seen at Lalo, Manawai, and Kapou. 

```{r}
tri_citizen <- list()

for(df_name in names(duplicate_ids)) {
  current_df <- duplicate_ids[[df_name]]
  unique_islands <- unique(current_df$Island)
  if(length(unique_islands) > 2) {
    tri_citizen[[df_name]] <- current_df
  }
}

specific_tri <- tri_citizen[['5579']]
specific_tri
```

##Sex distribution of the turtles that moved 

To find the sex distribution of the honu that moved, first I need to make a new list with moving turtles with only the observations that had sex information (which was over 90% of them). Then I looped through and made a count variable to keep track of how many male and female turtles were in the moving turtle data frame. Of the turtles that moved (and had sex information) 116 were females and 12 were males.  

```{r}
female_counts_move <- 0
male_counts_move <- 0

moving_turtles_w_sex <- list()

#This for loop filters out any dataframes that do not have either an M or F in their sex column 
for(df_name in names(moving_turtles)) {
  current_df <- moving_turtles[[df_name]]
  #gets rid of any rows in the data frame that have a U for unknown sex
  filtered_df <- current_df[current_df$Sex != "U", ]
  # Check if the current data frame has sex information (i.e. not NA)
  if (!any(is.na(filtered_df$Sex))) {
    # Add the data frame to the moving_turtles_w_sex list
    moving_turtles_w_sex[[df_name]] <- filtered_df
  }
}

#This for loop actually counts the instances of F and M
for(df_name in names(moving_turtles_w_sex)) {
  current_df <- moving_turtles_w_sex[[df_name]]

  if(!is.na(current_df[1, "Sex"]) && current_df[1, "Sex"] == "F") {
    female_counts_move = female_counts_move + 1
  }
  
  if(!is.na(current_df[1, "Sex"]) && current_df[1, "Sex"] == "M") {
    male_counts_move = male_counts_move + 1
  }
}

Sex <- c("Female", "Male")
Moving_Sums <- c(female_counts_move, male_counts_move)
sex_distribution_moving <- data.frame(Sex, Moving_Sums)

sex_distribution_moving

```

##Number of female vs males that we have tagged 

Now I need to figure out if the above female vs male movement metric is a result of a bias in our tagging effort, since we spend a lot of time tagging nesting females. So I need to get total female and male tagging counts from the original unique turtle data frame (using only turtles with sex information obviously so it won't be exact). This code is pretty much the same as the chunk above, just with a different list. Of the 9,246 unique tagged turtles with sex information over the history of the entire program, 7473 turtles were female and 1773 were male. 

```{r}
female_counts <- 0
male_counts <- 0

duplicate_ids_w_sex <- list()

for(df_name in names(duplicate_ids)) {
  current_df <- duplicate_ids[[df_name]]
  #gets rid of any rows in the dataframe that have a U for unknown sex
  filtered_df <- current_df[current_df$Sex != "U", ]
  # Check if the current dataframe has sex information (ie not NA)
  if (!any(is.na(filtered_df$Sex))) {
    # Add the dataframe to the duplicate_ids_w_sex list
    duplicate_ids_w_sex[[df_name]] <- filtered_df
  }
}

for(df_name in names(duplicate_ids_w_sex)) {
  current_df <- duplicate_ids_w_sex[[df_name]]

  if(!is.na(current_df[1, "Sex"]) && current_df[1, "Sex"] == "F") {
    female_counts = female_counts + 1
  }

  if(!is.na(current_df[1, "Sex"]) && current_df[1, "Sex"] == "M") {
    male_counts = male_counts + 1
  }
}

Sex <- c("Female", "Male")
Tagging_Sums <- c(female_counts, male_counts)
sex_distribution_total <- data.frame(Sex, Tagging_Sums)

sex_distribution_total
```

##Has a turtle ever been seen on different islands in the same year? 

Yes. 15 turtles have been seen at two islands within the same year. Most were seen at Lalo during the nesting season and at another northern island during the off season. 

```{r}
turtles_multiple_islands <- list()

for (df_name in names(moving_turtles)) {
  current_df <- moving_turtles[[df_name]]
  unique_years <- unique(current_df$Year)
  # will turn true if it finds a turtle that went to multiple islands
  multiple_islands_flag <- FALSE
  for (year in unique_years) {
    year_rows <- current_df[current_df$Year == year, ]
    unique_islands <- unique(year_rows$Island)
    if (length(unique_islands) > 1) {
      multiple_islands_flag <- TRUE
      break
    }
  }
  # Add the dataframe to the turtles_multiple_islands list if sightings in multiple islands were found
  if (multiple_islands_flag) {
    turtles_multiple_islands[[df_name]] <- current_df
  }
}

specific_mover <- turtles_multiple_islands[['20180506FF35EM']]
specific_mover
```

##New dataframe

This simplified list of data frames has just the first observation for each turtle for each year at a unique island. I want to use this to make it easier for me to track movements between the islands in code. So it gets rid of duplicate observations of the same turtle on the same island during the same year. You should still be able to track which direction the turtle is going based on the dates.  

```{r}
simplified_moving_turtles <- list()

for (df_name in names(moving_turtles)) {
  current_df <- moving_turtles[[df_name]]
  # Group the dataframe by Turtle_ID, Year, and Island, and keep the first observation for each group
  simplified_df <- current_df[!duplicated(current_df[, c("Turtle_ID", "Year", "Island")]), ]
  simplified_moving_turtles[[df_name]] <- simplified_df
}

```

##Lisi connections

Now I want to focus on just Lisianski and find out of the turtles that went to Lisi, what is the second (or third) island they went to?

```{r}
lisi_movements <- list()

for(df_name in names(simplified_moving_turtles)) {
  current_df <- simplified_moving_turtles[[df_name]]
  unique_islands <- unique(current_df$Island)
  for(island in unique_islands) {
    if(island == "LISI") {
      lisi_movements[[df_name]] <- current_df
    }
  }
}

##Getting rid of extra observations and simplifying it down to just which two (or three hello 5579) islands each turtle went to so that I'm not double counting islands
unique_lisi_movements <- list()

for(df_name in names(lisi_movements)) {
  current_df <- lisi_movements[[df_name]]
  unique_islands <- unique(current_df$Island)
  unique_island_df <- data.frame()
  for(island in unique_islands){
    island_observation <- current_df[current_df$Island == island, ][1, ]
    unique_island_df <- rbind(unique_island_df, island_observation)
  }
  unique_lisi_movements[[df_name]] <- unique_island_df
}

ffs_lisi_count <- 0 
ph_lisi_count <- 0
mid_lisi_count <- 0
lay_lisi_count <- 0 
kure_lisi_count <- 0
neck_lisi_count <- 0

for(df_name in names(unique_lisi_movements)){
  current_df <- unique_lisi_movements[[df_name]]
    # Iterate over each row in the current dataframe
  for (i in 1:nrow(current_df)) {
    # Get the island code of the current row
    island_code <- current_df[i, "Island"]
    
    # Increment the count for the corresponding island
    if (island_code == "FFS") {
      ffs_lisi_count <- ffs_lisi_count + 1
    } else if (island_code == "PEAR") {
      ph_lisi_count <- ph_lisi_count + 1
    } else if (island_code == "MIDW") {
      mid_lisi_count <- mid_lisi_count + 1
    } else if (island_code == "LAYS") {
      lay_lisi_count <- lay_lisi_count + 1
    } else if (island_code == "KURE") {
      kure_lisi_count <- kure_lisi_count + 1
    } else if (island_code == "NECK") {
      neck_lisi_count <- neck_lisi_count + 1
    }
  }
}

island_counts_lisi <- data.frame(
  Island = c("FFS", "PEAR", "MIDW", "LAYS", "KURE", "NECK"),
  Count = c(ffs_lisi_count, ph_lisi_count, mid_lisi_count, lay_lisi_count, kure_lisi_count, neck_lisi_count)
)
island_counts_lisi
```

##Midway Connections

Same thing but for Midway. 

```{r}
midway_movements <- list()

for(df_name in names(simplified_moving_turtles)) {
  current_df <- simplified_moving_turtles[[df_name]]
  unique_islands <- unique(current_df$Island)
  for(island in unique_islands) {
    if(island == "MIDW") {
      midway_movements[[df_name]] <- current_df
    }
  }
}

unique_midway_movements <- list()

for(df_name in names(midway_movements)) {
  current_df <- midway_movements[[df_name]]
  unique_islands <- unique(current_df$Island)
  unique_island_df <- data.frame()
  for(island in unique_islands){
    island_observation <- current_df[current_df$Island == island, ][1, ]
    unique_island_df <- rbind(unique_island_df, island_observation)
  }
  unique_midway_movements[[df_name]] <- unique_island_df
}

ffs_mid_count <- 0 
ph_mid_count <- 0
lisi_mid_count <- 0
lay_mid_count <- 0 
kure_mid_count <- 0
neck_mid_count <- 0

for(df_name in names(unique_midway_movements)){
  current_df <- unique_midway_movements[[df_name]]
    # Iterate over each row in the current dataframe
  for (i in 1:nrow(current_df)) {
    # Get the island code of the current row
    island_code <- current_df[i, "Island"]
    
    # Increment the count for the corresponding island
    if (island_code == "FFS") {
      ffs_mid_count <- ffs_mid_count + 1
    } else if (island_code == "PEAR") {
      ph_mid_count <- ph_mid_count + 1
    } else if (island_code == "LISI") {
      lisi_mid_count <- lisi_mid_count + 1
    } else if (island_code == "LAYS") {
      lay_mid_count <- lay_mid_count + 1
    } else if (island_code == "KURE") {
      kure_mid_count <- kure_mid_count + 1
    } else if (island_code == "NECK") {
      neck_mid_count <- neck_mid_count + 1
    }
  }
}

# Create a new dataframe with the island counts
island_counts_mid <- data.frame(
  Island = c("FFS", "PEAR", "LISI", "LAYS", "KURE", "NECK"),
  Count = c(ffs_mid_count, ph_mid_count, lisi_mid_count, lay_mid_count, kure_mid_count, neck_mid_count)
)

# Display the island_counts dataframe
island_counts_mid
```

##Pearl and Hermes Connections

Same thing but for Manawai. 

```{r}
ph_movements <- list()

for(df_name in names(simplified_moving_turtles)) {
  current_df <- simplified_moving_turtles[[df_name]]
  unique_islands <- unique(current_df$Island)
  for(island in unique_islands) {
    if(island == "PEAR") {
      ph_movements[[df_name]] <- current_df
    }
  }
}

unique_ph_movements <- list()

for(df_name in names(ph_movements)) {
  current_df <- ph_movements[[df_name]]
  unique_islands <- unique(current_df$Island)
  unique_island_df <- data.frame()
  for(island in unique_islands){
    island_observation <- current_df[current_df$Island == island, ][1, ]
    unique_island_df <- rbind(unique_island_df, island_observation)
  }
  unique_ph_movements[[df_name]] <- unique_island_df
}

ffs_ph_count <- 0 
mid_ph_count <- 0
lisi_ph_count <- 0
lay_ph_count <- 0 
kure_ph_count <- 0
neck_ph_count <- 0

for(df_name in names(unique_ph_movements)){
  current_df <- unique_ph_movements[[df_name]]
    # Iterate over each row in the current dataframe
  for (i in 1:nrow(current_df)) {
    # Get the island code of the current row
    island_code <- current_df[i, "Island"]
    
    # Increment the count for the corresponding island
    if (island_code == "FFS") {
      ffs_ph_count <- ffs_ph_count + 1
    } else if (island_code == "MIDW") {
      mid_ph_count <- mid_ph_count + 1
    } else if (island_code == "LISI") {
      lisi_ph_count <- lisi_ph_count + 1
    } else if (island_code == "LAYS") {
      lay_ph_count <- lay_ph_count + 1
    } else if (island_code == "KURE") {
      kure_ph_count <- kure_ph_count + 1
    } else if (island_code == "NECK") {
      neck_ph_count <- neck_ph_count + 1
    }
  }
}

# Create a new dataframe with the island counts
island_counts_ph <- data.frame(
  Island = c("FFS", "MIDW", "LISI", "LAYS", "KURE", "NECK"),
  Count = c(ffs_ph_count, mid_ph_count, lisi_ph_count, lay_ph_count, kure_ph_count, neck_ph_count)
)

# Display the island_counts dataframe
island_counts_ph
```

##Laysan Connections

Same thing but for Kamole. 

```{r}
lay_movements <- list()

for(df_name in names(simplified_moving_turtles)) {
  current_df <- simplified_moving_turtles[[df_name]]
  unique_islands <- unique(current_df$Island)
  for(island in unique_islands) {
    if(island == "LAYS") {
      lay_movements[[df_name]] <- current_df
    }
  }
}

unique_lay_movements <- list()

for(df_name in names(lay_movements)) {
  current_df <- lay_movements[[df_name]]
  unique_islands <- unique(current_df$Island)
  unique_island_df <- data.frame()
  for(island in unique_islands){
    island_observation <- current_df[current_df$Island == island, ][1, ]
    unique_island_df <- rbind(unique_island_df, island_observation)
  }
  unique_lay_movements[[df_name]] <- unique_island_df
}

ffs_lay_count <- 0 
ph_lay_count <- 0
lisi_lay_count <- 0
mid_lay_count <- 0 
kure_lay_count <- 0
neck_lay_count <- 0

for(df_name in names(unique_lay_movements)){
  current_df <- unique_lay_movements[[df_name]]
    # Iterate over each row in the current dataframe
  for (i in 1:nrow(current_df)) {
    # Get the island code of the current row
    island_code <- current_df[i, "Island"]
    
    # Increment the count for the corresponding island
    if (island_code == "FFS") {
      ffs_lay_count <- ffs_lay_count + 1
    } else if (island_code == "PEAR") {
      ph_lay_count <- ph_lay_count + 1
    } else if (island_code == "LISI") {
      lisi_lay_count <- lisi_lay_count + 1
    } else if (island_code == "MIDW") {
      mid_lay_count <- mid_lay_count + 1
    } else if (island_code == "KURE") {
      kure_lay_count <- kure_lay_count + 1
    } else if (island_code == "NECK") {
      neck_lay_count <- neck_lay_count + 1
    }
  }
}

# Create a new dataframe with the island counts
island_counts_lay <- data.frame(
  Island = c("FFS", "PEAR", "LISI", "MIDW", "KURE", "NECK"),
  Count = c(ffs_lay_count, ph_lay_count, lisi_lay_count, mid_lay_count, kure_lay_count, neck_lay_count)
)

# Display the island_counts dataframe
island_counts_lay
```

##Necker movements

Same thing but for Mokumanamana.

```{r}
neck_movements <- list()

for(df_name in names(simplified_moving_turtles)) {
  current_df <- simplified_moving_turtles[[df_name]]
  unique_islands <- unique(current_df$Island)
  for(island in unique_islands) {
    if(island == "NECK") {
      neck_movements[[df_name]] <- current_df
    }
  }
}

unique_neck_movements <- list()

for(df_name in names(neck_movements)) {
  current_df <- neck_movements[[df_name]]
  unique_islands <- unique(current_df$Island)
  unique_island_df <- data.frame()
  for(island in unique_islands){
    island_observation <- current_df[current_df$Island == island, ][1, ]
    unique_island_df <- rbind(unique_island_df, island_observation)
  }
  unique_neck_movements[[df_name]] <- unique_island_df
}

ffs_neck_count <- 0 
ph_neck_count <- 0
lisi_neck_count <- 0
mid_neck_count <- 0 
kure_neck_count <- 0
lay_neck_count <- 0

for(df_name in names(unique_neck_movements)){
  current_df <- unique_neck_movements[[df_name]]
    # Iterate over each row in the current dataframe
  for (i in 1:nrow(current_df)) {
    # Get the island code of the current row
    island_code <- current_df[i, "Island"]
    
    # Increment the count for the corresponding island
    if (island_code == "FFS") {
      ffs_neck_count <- ffs_neck_count + 1
    } else if (island_code == "PEAR") {
      ph_neck_count <- ph_neck_count + 1
    } else if (island_code == "LISI") {
      lisi_neck_count <- lisi_neck_count + 1
    } else if (island_code == "MIDW") {
      mid_neck_count <- mid_neck_count + 1
    } else if (island_code == "KURE") {
      kure_neck_count <- kure_neck_count + 1
    } else if (island_code == "LAYS") {
      lay_neck_count <- lay_neck_count + 1
    }
  }
}

# Create a new dataframe with the island counts
island_counts_neck <- data.frame(
  Island = c("FFS", "PEAR", "LISI", "MIDW", "KURE", "LAYS"),
  Count = c(ffs_neck_count, ph_neck_count, lisi_neck_count, mid_neck_count, kure_neck_count, lay_neck_count)
)

# Display the island_counts dataframe
island_counts_neck
```

##Kure connections

Same thing but for Holaniku. 

```{r}
kure_movements <- list()

for(df_name in names(simplified_moving_turtles)) {
  current_df <- simplified_moving_turtles[[df_name]]
  unique_islands <- unique(current_df$Island)
  for(island in unique_islands) {
    if(island == "KURE") {
      kure_movements[[df_name]] <- current_df
    }
  }
}

unique_kure_movements <- list()

for(df_name in names(kure_movements)) {
  current_df <- kure_movements[[df_name]]
  unique_islands <- unique(current_df$Island)
  unique_island_df <- data.frame()
  for(island in unique_islands){
    island_observation <- current_df[current_df$Island == island, ][1, ]
    unique_island_df <- rbind(unique_island_df, island_observation)
  }
  unique_kure_movements[[df_name]] <- unique_island_df
}

ffs_kure_count <- 0 
ph_kure_count <- 0
lisi_kure_count <- 0
mid_kure_count <- 0 
neck_kure_count <- 0
lay_kure_count <- 0

for(df_name in names(unique_kure_movements)){
  current_df <- unique_kure_movements[[df_name]]
    # Iterate over each row in the current dataframe
  for (i in 1:nrow(current_df)) {
    # Get the island code of the current row
    island_code <- current_df[i, "Island"]
    
    # Increment the count for the corresponding island
    if (island_code == "FFS") {
      ffs_kure_count <- ffs_kure_count + 1
    } else if (island_code == "PEAR") {
      ph_kure_count <- ph_kure_count + 1
    } else if (island_code == "LISI") {
      lisi_kure_count <- lisi_kure_count + 1
    } else if (island_code == "MIDW") {
      mid_kure_count <- mid_kure_count + 1
    } else if (island_code == "NECK") {
      neck_kure_count <- neck_kure_count + 1
    } else if (island_code == "LAYS") {
      lay_kure_count <- lay_kure_count + 1
    }
  }
}

# Create a new dataframe with the island counts
island_counts_kure <- data.frame(
  Island = c("FFS", "PEAR", "LISI", "MIDW", "NECK", "LAYS"),
  Count = c(ffs_kure_count, ph_kure_count, lisi_kure_count, mid_kure_count, neck_kure_count, lay_kure_count)
)

# Display the island_counts dataframe
island_counts_kure
```


##Resights of only Lalo nesters

This is to explore how many of the moving turtles were confirmed Lalo nesters. Of the 116 moving females, 94 of them were nesters at Lalo. 

```{r}
# New list of data frames for each of the moving turtles who were nesters at Lalo. 
moving_lalo_nesters <- list()

for(df_name in names(moving_turtles)) {
  current_df <- moving_turtles[[df_name]]
  for(i in 1:nrow(current_df)){
    island_code <- current_df[i, "Island"]
    event_code <- current_df[i, "Event Type"]
    if(island_code == "FFS" & event_code == "Nesting"){
      moving_lalo_nesters[[df_name]] <- current_df
    }
  }
}

specific_lalo_mover <- moving_lalo_nesters[['1260']]
specific_lalo_mover
```

##Where are the nesting females going? 

Now I'm curious about what islands these nesting females are going to...

Seems like they're mostly going to Manawai and Midway, which makes sense since they have the most total connections to Lalo (52 between Lalo and Manawai and 35 between Lalo and Midway). 

```{r}
unique_moving_lalo_nesters_movements <- list()

##Simplifying the dataframes down to unique observations at each island to make them easier to count
for(df_name in names(moving_lalo_nesters)) {
  current_df <- moving_lalo_nesters[[df_name]]
  unique_islands <- unique(current_df$Island)
  unique_island_df <- data.frame()
  for(island in unique_islands){
    island_observation <- current_df[current_df$Island == island, ][1, ]
    unique_island_df <- rbind(unique_island_df, island_observation)
  }
  unique_moving_lalo_nesters_movements[[df_name]] <- unique_island_df
}

lalo_nesters_at_lay <- 0
lalo_nesters_at_lisi <- 0
lalo_nesters_at_ph <- 0
lalo_nesters_at_mid <- 0
lalo_nesters_at_kure <- 0
lalo_nesters_at_necker <- 0

for(df_name in names(unique_moving_lalo_nesters_movements)) {
  current_df <- unique_moving_lalo_nesters_movements[[df_name]]
  unique_islands <- unique(current_df$Island)
  for(island in unique_islands){
    if(island == "LAYS"){
      lalo_nesters_at_lay = lalo_nesters_at_lay + 1
    }
    if(island == "LISI"){
      lalo_nesters_at_lisi = lalo_nesters_at_lisi + 1
    }
    if(island == "PEAR"){
      lalo_nesters_at_ph = lalo_nesters_at_ph +1
    }
    if(island == "MIDW"){
      lalo_nesters_at_mid = lalo_nesters_at_mid +1
    }
    if(island == "NECK"){
      lalo_nesters_at_necker = lalo_nesters_at_necker +1
    }
    if(island == "KURE"){
      lalo_nesters_at_kure = lalo_nesters_at_kure +1
    }
  }
}


lalo_nesters_paired_islands <- data.frame(
  Island = c("Necker", "Laysan", "Lisianski", "Pearl & Hermes", "Midway", "Kure"),
  Count = c(lalo_nesters_at_necker, lalo_nesters_at_lay, lalo_nesters_at_lisi, lalo_nesters_at_ph, lalo_nesters_at_mid, lalo_nesters_at_kure)
)
lalo_nesters_paired_islands

```

