---
title: "Signficance Testing"
output: html_document
date: "2023-06-29"
---

##Libraries

```{r setup, include=FALSE}
library(here)
library(tidyverse)
library(readxl)
library(lme4)
library(multcomp)
library(dplyr)
```

##Data

```{r}
dat <- SizeData
Lalo_dat <- combined_lalo

Lalo_dat <- Lalo_dat[!duplicated(Lalo_dat[c("Turtle ID", "Year")]), ]

```

##Making new dataframes

```{r}
append_Lalo_F <- Lalo_dat %>%
  dplyr::filter(Sex == "F") %>%
  dplyr::mutate(CCL_com = coalesce(CCL, Est_CCL)) %>%
  dplyr::select(Island, CCL_com, Year)

nwhi_females <- dat %>%
  dplyr::filter(Sex == "F", Island != "NECK") %>%
  dplyr::mutate(CCL_com = coalesce(CCL, Est_CCL)) %>%
  dplyr::select(Island, CCL_com, Year) 

append_Lalo_M <- Lalo_dat %>%
  dplyr::filter(Sex == "M") %>%
  dplyr::mutate(CCL_com = coalesce(CCL, Est_CCL)) %>%
  dplyr::select(Island, CCL_com, Year)

nwhi_males <- dat %>%
  dplyr::filter(Sex == "M", Island != "NECK") %>%
  dplyr::mutate(CCL_com = coalesce(CCL, Est_CCL)) %>%
  dplyr::select(Island, CCL_com, Year)

##data frame of all raw data of females from all islands
all_females <- bind_rows(dplyr::select(nwhi_females, Island, Year, CCL_com), dplyr::select(append_Lalo_F, Island, Year, CCL_com)) 

##data frame of all raw data of males from all islands
all_males <- bind_rows(dplyr::select(nwhi_males, Island, Year, CCL_com), dplyr::select(append_Lalo_M, Island, Year, CCL_com)) 

```

##Fitting Linear Models

```{r}
all_females$Year <- as.numeric(all_females$Year)
all_males$Year <- as.numeric(all_males$Year)

#model for females
females_lm <- lm(CCL_com ~ (0+Island) + Year:(0+Island), data = all_females)
summary(females_lm)

#model for males
males_lm <- lm(CCL_com ~ (0+Island) + Year:(0+Island), data = all_males)

```

##Contrast test to see if there are significant differences >:)

No signficant differences though in the slopes between the different islands :( 

```{r}
##Female contrast testing
female_K <- matrix(c(0, 0, 0, 0, 0, 1, -1, 0, 0, 0,
                     0, 0, 0, 0, 0, 1, 0, -1, 0, 0, 
                     0, 0, 0, 0, 0, 1, 0, 0, -1, 0,
                     0, 0, 0, 0, 0, 1, 0, 0, 0, -1,
                     0, 0, 0, 0, 0, 0, 1, -1, 0, 0,
                     0, 0, 0, 0, 0, 0, 1, 0, -1, 0,
                     0, 0, 0, 0, 0, 0, 1, 0, 0, -1,
                     0, 0, 0, 0, 0, 0, 0, 1, -1, 0,
                     0, 0, 0, 0, 0, 0, 0, 1, 0, -1,
                     0, 0, 0, 0, 0, 0, 0, 0, 1, -1), nrow = 10, byrow = TRUE)
female_t <- glht(females_lm, linfct = female_K)
summary(female_t)

##Male contrast testing
male_K <- matrix(c(0, 0, 0, 0, 0, 1, -1, 0, 0, 0,
                     0, 0, 0, 0, 0, 1, 0, -1, 0, 0, 
                     0, 0, 0, 0, 0, 1, 0, 0, -1, 0,
                     0, 0, 0, 0, 0, 1, 0, 0, 0, -1,
                     0, 0, 0, 0, 0, 0, 1, -1, 0, 0,
                     0, 0, 0, 0, 0, 0, 1, 0, -1, 0,
                     0, 0, 0, 0, 0, 0, 1, 0, 0, -1,
                     0, 0, 0, 0, 0, 0, 0, 1, -1, 0,
                     0, 0, 0, 0, 0, 0, 0, 1, 0, -1,
                     0, 0, 0, 0, 0, 0, 0, 0, 1, -1), nrow = 10, byrow = TRUE)
male_t <- glht(males_lm, linfct = male_K)
summary(male_t)

```

##Testing to see if there are significant differences between males and females at the same island

##new raw data dataframes for each Island with both sexes 

Now I'm creating new dataframes with just the raw data from each island with both sexes to determine if there is a significant difference in the sizes of males and females for all years

```{r}
Laysan_raw <- dat %>%
  filter(Island == "LAYS") %>%
  dplyr::mutate(CCL_com = coalesce(CCL, Est_CCL)) %>%
  dplyr::select(Island, Sex, CCL_com, Year)

Lisi_raw <- dat %>%
  filter(Island == "LISI") %>%
  dplyr::mutate(CCL_com = coalesce(CCL, Est_CCL)) %>%
  dplyr::select(Island, Sex, CCL_com, Year) %>%
  na.omit()

Midway_raw <- dat %>%
  filter(Island == "MIDW") %>%
  dplyr::mutate(CCL_com = coalesce(CCL, Est_CCL)) %>%
  dplyr::select(Island, Sex, CCL_com, Year)

PH_raw <- dat %>%
  filter(Island == "PEAR") %>%
  dplyr::mutate(CCL_com = coalesce(CCL, Est_CCL)) %>%
  dplyr::select(Island, Sex, CCL_com, Year)

Lalo_raw <- Lalo_dat %>%
  dplyr::mutate(CCL_com = coalesce(CCL, Est_CCL)) %>%
  dplyr::select(Island, Sex, CCL_com, Year)
```

##Linear models to test significance 

```{r}
Laysan_lm <- lm(CCL_com ~ Sex, Laysan_raw)
summary(Laysan_lm)

Lisi_lm <- lm(CCL_com ~ Sex, Lisi_raw)
summary(Lisi_lm)

Midway_lm <- lm(CCL_com ~ Sex, Midway_raw)
summary(Midway_lm)

PH_lm <- lm(CCL_com ~ Sex, PH_raw)
summary(PH_lm)

Lalo_lm <- lm(CCL_com ~ Sex, Lalo_raw)
summary(Lalo_lm)
```

##Testing difference between NHI and Lalo average CCL for males and females 

```{r}
nwhi_females2 <- nwhi_females %>%
  dplyr::select(CCL_com, Year) %>%
  dplyr::mutate(Island = "NWHI")

nwhi_males2 <- nwhi_males %>%
  dplyr::select(CCL_com, Year) %>%
  dplyr::mutate(Island = "NWHI")

comp_females <- bind_rows(dplyr::select(nwhi_females2, Island, Year, CCL_com), dplyr::select(append_Lalo_F, Island, Year, CCL_com)) 

comp_males <- bind_rows(dplyr::select(nwhi_males2, Island, Year, CCL_com), dplyr::select(append_Lalo_M, Island, Year, CCL_com)) 

females_lm <- lm(CCL_com ~ Island, comp_females)
summary(females_lm)

males_lm <- lm(CCL_com ~ Island, comp_males)
summary(males_lm)

```


